### 2023-04-04 blog.qmd
Create blog.qmd

---
title: "Biomarker One Day"
listing:
  contents: posts
  sort: "date desc"
  type: default
  categories: true
  sort-ui: true
  filter-ui: true
  fields: [date, title, reading-time, description, categories]
page-layout: full
title-block-banner: false
---

# Series

##### [rtables-series](rtables-series.html)
This series contains a great deal of tips, tricks and packages 
that you can use to level up your rtables game.

### 2023-04-15 blog.qmd
Update blog.qmd

@@
 ---
-title: "Biomarker One Day"
+title: "One Day"
 listing:
@@
 This series contains a great deal of tips, tricks and packages
-that you can use to level up your rtables game.
+that you can use to level up your `rtables` game.
+
+##### [manipulate-series](manipulate-series.html)
+This series contains a great deal of tips, tricks of `dplyr`

### 2023-04-15 posts/rtables/customize_cell.qmd
Create customize_cell.qmd

---
title: 'Customize the format of rcell'
date: '2023-04-06'
author: Qing Zou
categories: ['R', 'rtables']
description: 'Construct a cell value and associate formatting'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

```{r rcell0}
library(rtables)
library(formatters)
```

## rcell

__Description__

Construct a cell value and associate formatting, labeling, indenting, and column spanning information with it.

__Usage__

```{r rcell1}
#| echo: true
#| eval: false
rcell(
x,
format = NULL,
colspan = 1L,
rcell 103
label = NULL,
indent_mod = NULL,
footnotes = NULL,
align = NULL,
format_na_str = NULL
)
```

__Arguments__

  |
 -|:--
 __x__| ANY. Cell value
 __format__| if FUN does not return a formatted rcell then the format is applied
 __...__|  further arguments passed to rcell

### Real case

If we have a calculation `1/4*100`, the result `25` and the decimal is `2`.
the avaiable format in rtables is `xx.xx`

```{r rcell2}
#| echo: true
#| eval: true
rcell(
  x = 25,
  format = 'xx.xx'
)
```

What if we want a format like `1 (25.00)`? how can we customize the format in rcell as expected?
Let's go through the function of `sprintf` firstly. the `sprintf` can returns a character vector containing a formatted combination of text and variable values.

1. print the `25` to `25.00`

```{r rcell3}
#| echo: true
#| eval: true
sprintf("%.2f", 25)
```
For the given format, `%`: it's like a place holder for the passed value, `f`: Double precision value, see more on [sprintf](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/sprintf)

2. update the format and passed value

```{r rcell4}
#| echo: true
#| eval: true
sprintf("%.0f %.2f", 1, 25)
```

```{r rcell5}
#| echo: true
#| eval: true
sprintf("%.0f (%.2f)", 1, 25)
```

---

In order to apply this format on other values, can we separate the formatting `%.0f (%.2f)` and the passed value `1` and `25`?

there are two functions from package `formatters`

`sprintf_format`: Specify text format

`format_value`: Converts a value into a string using the format information

```{r rcell6}
#| echo: true
#| eval: true
fmtfun <- formatters::sprintf_format("%.0f (%.2f)")

format_value(x = c(1, 25),
             format = fmtfun)
```

As for the `rcell` and  its argument `x` and `format`, we can think it is a wrapper around the `format_value`

```{r rcell7}
#| echo: true
#| eval: true
rcell(
  x = c(1, 25),
  format = formatters::sprintf_format("%.0f (%.2f)")
)
```

Then, we can also easily customize the result of `1/4*100` as `1/4 (25.00)`

```{r rcell8}
#| echo: true
#| eval: true
rcell(
  x = c(1, 4, 25),
  format = formatters::sprintf_format("%.0f/%.0f (%.2f)")
)
```

### Create format helper function

we can further more wrap this pair of funcions into two helper function for better usage.

```{r rcell9}
#| echo: true
#| eval: true

#' a pair of functions sprintf_format and format_values used for rcell
#'
#' @param n Unique number of events for the current category
#' @param N Total number of events under the split column
#' @param decimal Number of decimal to display
#' @param type to decide if the denominator n/N(%) in the result
#' @keywords sprintf_format
#'
#' @return
sprintf_format_helper <- function(n, N, decimal = 1, type = "n(%)") {
  showN <- NULL
  if (type == "n/N(%)"){
    showN <- "/%.0f"
  }

  if (n == 0) {
    fmt <- sprintf_format(paste0("%.0f", showN))
  } else if (n == N) {
    # if pct == 100%, then show (100) instead of (100.0)
    fmt <- sprintf_format(paste0("%.0f", showN, " ", "(%.0f)"))
  } else {
    fmt <- sprintf_format(paste0("%.0f", showN, " (%.", decimal, "f)"))
  }

  return(fmt)
}

#' a pair of functions sprintf_format and format_values used for rcell
#'
#' @param n Unique number of events for the current category
#' @param N Total number of events under the split column
#' @param type to decide if the denominator n/N(%) in the result
#'
#' @keywords format_values
#' @example
#' rcell(x = format_values_helper(1, 10),
#'       format = sprintf_format_helper(1, 10))
#'
#' rcell(x = format_values_helper(1, 10, type = "n/N(%)"),
#'       format = sprintf_format_helper(1, 10, type = "n/N(%)"))
#'
#' rcell(x = format_values_helper(10, 10),
#'       format = sprintf_format_helper(10, 10))
#'
#' @return
format_values_helper <- function(n, N, type = "n(%)") {
  showN <- NULL
  if (type == "n/N(%)"){
    showN <- N
  }
  if (n == 0) {
    value <- c(0, showN)
  } else {
    value <- c(n, showN, n / N * 100)
  }

  return(value)
}
```

then, we can apply various formatter during printing values in `rcell`.

```{r rcell10}
#| echo: true
#| eval: true
rcell(x = format_values_helper(1, 10),
      format = sprintf_format_helper(1, 10))
```

```{r rcell11}
#| echo: true
#| eval: true

rcell(x = format_values_helper(1, 10, type = "n/N(%)"),
      format = sprintf_format_helper(1, 10, type = "n/N(%)"))

```

```{r rcell12}
#| echo: true
#| eval: true

rcell(x = format_values_helper(10, 10),
      format = sprintf_format_helper(10, 10))
```

```{r rcell13}
#| echo: true
#| eval: true

rcell(x = format_values_helper(0, 10, type = "n/N(%)"),
      format = sprintf_format_helper(0, 10, type = "n/N(%)"))
```

### 2023-04-15 posts/rtables/multiple_levels_calculation_1.qmd
Create multiple_levels_calculation_1.qmd

---
title: 'Mutiple level calculation 1'
date: '2023-04-07'
author: Qing Zou
categories: ['R', 'rtables']
description: 'Explore how to handle the mutiple calculation by rtables'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

```{r rcell3sd45}
#| echo: false
#| eval: true

#' a pair of functions sprintf_format and format_values used for rcell
#'
#' @param n Unique number of events for the current category
#' @param N Total number of events under the split column
#' @param decimal Number of decimal to display
#' @param type to decide if the denominator n/N(%) in the result
#' @keywords sprintf_format
#'
#' @return
sprintf_format_helper <- function(n, N, decimal = 1, type = "n(%)") {
  showN <- NULL
  if (type == "n/N(%)"){
    showN <- "/%.0f"
  }

  if (n == 0) {
    fmt <- sprintf_format(paste0("%.0f", showN))
  } else if (n == N) {
    # if pct == 100%, then show (100) instead of (100.0)
    fmt <- sprintf_format(paste0("%.0f", showN, " ", "(%.0f)"))
  } else {
    fmt <- sprintf_format(paste0("%.0f", showN, " (%.", decimal, "f)"))
  }

  return(fmt)
}

#' a pair of functions sprintf_format and format_values used for rcell
#'
#' @param n Unique number of events for the current category
#' @param N Total number of events under the split column
#' @param type to decide if the denominator n/N(%) in the result
#'
#' @keywords format_values
#' @example
#' rcell(x = format_values_helper(1, 10),
#'       format = sprintf_format_helper(1, 10))
#'
#' rcell(x = format_values_helper(1, 10, type = "n/N(%)"),
#'       format = sprintf_format_helper(1, 10, type = "n/N(%)"))
#'
#' rcell(x = format_values_helper(10, 10),
#'       format = sprintf_format_helper(10, 10))
#'
#' @return
format_values_helper <- function(n, N, type = "n(%)") {
  showN <- NULL
  if (type == "n/N(%)"){
    showN <- N
  }
  if (n == 0) {
    value <- c(0, showN)
  } else {
    value <- c(n, showN, n / N * 100)
  }

  return(value)
}

#' Display rtables in html
#'
#' @param table a rtables object
#' @param to_html export to HTML. Default TRUE
#'
#' @return a html table
#' @export
table_to_html <- function(table, to_html = TRUE) {
  stopifnot(!is.null(table), class(table) %in% c("TableTree", "ElementaryTable"))

  extra_css <- "
  .tabwid {overflow-x: scroll; display: block;}
  .tabwid table td p {min-width: 170px !important;}
  .tabwid table td:first-child > p {min-width: 200px !important;}
  .tabwid table tr:hover {background-color: #ffff99 !important;}
  "
  flextable::set_flextable_defaults(
    font.family = ifelse(to_html, "Source Sans Pro", "Verdana"),
    extra_css = extra_css
  )

  ft <- rtables::tt_to_flextable(table) %>%
    flextable::border_remove() %>%
    flextable::bold(bold = TRUE, part = "header") %>%
    flextable::hline_top(part = "body") %>%
    flextable::hline_bottom(part = "body") %>%
    flextable::autofit()

  if (to_html) {
    flextable::htmltools_value(ft)
  } else {
    ft
  }
}
```

```{r cal1}
library(rtables)
library(formatters)
library(dplyr)
```

## Multiple lelvels calculation

`Multipile levels` calculation for categorical variable is the most common cases in Table, e.g. `PCSA`, `Exposure tables`.

In this section, we will explore how to produce these output in `rtables`.

Below is the test data:

```{r cal2}
#| echo: true
#| eval: true
df <- data.frame(SUBJID = c(rep('S1', 5), rep("S2", 4), rep("S3", 2)),
                 ASEQ = c(1:5, 1:4, 1:2),
                 GROUP = factor(c(rep('A', 5), rep("B", 4), rep('A', 2)), levels = c("A", "B")),
                 PARAMCD = c('TEST_1', 'TEST_1', 'TEST_1', 'TEST_1', 'TEST_2', 'TEST_1', 'TEST_1', 'TEST_2', 'TEST_2', 'TEST_1', 'TEST_2'),
                 CRIT = c('LEVEL_2_1','LEVEL_2_1', 'LEVEL_2_2', NA_character_, 'LEVEL_2_a', 'LEVEL_2_1', NA_character_, 'LEVEL_2_a', 'LEVEL_2_b', NA_character_, 'LEVEL_2_b'))
```
```{r cal2sd }
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(df, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

Now, we need to take the `PARAMCD` as 1st level, and the `CRIT` as 2nd level to calculate the frequency of each `TEST`, it will be a table like PCSA.

# Method 1

the `rtables::analyze()` will firstly be used in this case.

To use this function, the data should be transformed from `long strcture(BDS)` to `wide`.

Steps:

1. take action on setting up for the `missing value` and the `factor`:

 - fill the missing value as `Missing_value_flag`, which will be subsequently handled in `rtables::analyze()`
 - use function `tidyr::pivot_wider` implement the transformation, the new columns are used as the analyzed variables `vars` of `rtables::analyze()`
 - `factor` the values by given level values, the levels is the format of output

```{r cal2df}
#| echo: true
#| eval: true

#
df <- df %>% mutate(across(CRIT, ~ifelse(.=="" | is.na(.), "Missing_value_flag", as.character(.))))

df1 <- df %>% tidyr::pivot_wider(id_cols = c("SUBJID", 'ASEQ', 'GROUP'),
                                 names_from = "PARAMCD",values_from = "CRIT")

#factor the data by mock
fmt <- list(
  'TEST_1' = c('LEVEL_2_1' = '1 of level 2',
               'LEVEL_2_2' = '2 of level 2',
               'LEVEL_2_3' = '3 of level 2'),
  "TEST_2" = c('LEVEL_2_a' = 'a of level 2',
              'LEVEL_2_b' = 'b of level 2',
              'LEVEL_2_c' = 'c of level 2')
)

fmt <- purrr::map(fmt, ~append(., c("Missing_value_flag" = "Missing_value_flag")))

df1 <- df1 %>%
  mutate(across(paste0('TEST_', 1:2), ~factor(fmt[[cur_column()]][.x],
                                              levels = c(fmt[[cur_column()]])), .names = "{.col}_fmt"))
```
```{r cal2g }
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(df1, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

2. The dataframe `df1` is ready for output. now, the `layout` and `key values` of table will be built in the `rtables::analyze()`.

 - the header and overall column are generated by  `split_cols_by` and `add_overall_col`
 - in the `rtables::analyze()`:
   - `vars` and `var_labels` are the variable names and labels
   - in the `afun`, we create an analysis function, in which the arguments `df`, `.N_col`, `.var`, `.N_row` are created by rtables itself. the other aruments are customized will be passed by `extral_args`
     - the `df` is current available(`non NA`) data with splited group, that's why we firstly fill the missing data `Missing_value_flag`. By filling
       a non-NA value, we can handle the missing value as users expected.
     - you will see the customized arguments `missing_rm` and `missing_label`, if user hope hide or remove the missing values.
     - the `n_unique_keys` and `N_unique_keys` indicate that when counting the number of n/N, which records should be used.
     - `add_Number` means whether add an row `Number: N`
     - All the values of these customized arguments are passed by the `extra_args`, it's the greatest way to add the flexiabiliy for the tabulation.

```{r cal2as}
#| echo: true
#| eval: true

lyt <- basic_table() %>%
  split_cols_by("GROUP") %>%
  add_overall_col("All") %>%
  rtables::analyze(vars = c('TEST_1_fmt', 'TEST_2_fmt'),
                   var_labels = c('This is test 1', 'This is test 2'),
                   afun = function(df, .N_col, .var, .N_row,
                                  format = 'n(%)',
                                  decimal = 1,
                                  n_unique_keys = NULL,
                                  N_unique_keys = NULL,
                                  show_N_item = FALSE,
                                  add_Number = FALSE,
                                  missing_rm = TRUE,
                                  missing_label =NULL,
                                  ...) {

                    if(missing_rm){
                      df <- df %>% filter(!(!!as.name(.var) %in% "Missing_value_flag"))
                    }

                    df_n <- df
                    if(!is.null(n_unique_keys)){
                      df_n <- df %>% distinct(across(c(n_unique_keys, .var)), .keep_all = TRUE)
                    }

                    x <- df_n[[.var]]

                    if(!is.null(N_unique_keys)){
                      cus_N_col <- df %>% distinct(across(N_unique_keys)) %>% nrow()
                    }else{
                      cus_N_col <- .N_col
                    }

                    lst_body <- lapply(
                      as.list(table(x)),
                      function(xi) {
                        rcell(format_values_helper(xi, cus_N_col, type = format),
                              format = sprintf_format_helper(xi, cus_N_col, 1, type = format),
                              indent_mod = 1L
                        )
                      }
                    )

                    if(is.null(missing_label)){
                      if("Missing_value_flag" %in% names(lst_body)){
                        lst_body <- purrr::list_modify(lst_body, "Missing_value_flag" = NULL)
                      }
                    }else{
                      lst_body <- lst_body %>%
                        setNames(gsub("Missing_value_flag", missing_label, names(.)))

                    }

                    if(add_Number){
                      lst_body <- append(setNames(list(rcell(cus_N_col,
                                                            format = "xx",
                                                            indent_mod = 1L)),
                                                  'Number'), lst_body)
                    }
                    in_rows(
                      .list = c(
                        lst_body,
                        # add empty row, in .list obj
                        " " = rcell("", format = "xx")
                      )
                    )
                  },
          extra_args = list(format = 'n/N(%)',
                            n_unique_keys = c("SUBJID"),
                            N_unique_keys = c("SUBJID"),
                            add_Number = FALSE,
                            missing_rm = FALSE,
                            missing_label = NULL)
          # ,
          # show_labels = "hidden"
  )
```
3. finally, fit the `layout` and `df1` into the `rtables::build_table`

```{r cal2ad}
#| echo: true
#| eval: true
col_counts_df <- df1 %>%
  distinct(SUBJID, GROUP)
col_counts <- table(col_counts_df$GROUP)
col_counts <- c(col_counts, sum(col_counts))

tbl <- lyt %>% rtables::build_table(df1, col_counts = col_counts)
```

```{r cap1}
#| echo: false
#| eval: true
my_caption <- ''
```

```{r}
#| echo: false
#| eval: true
#| message: FALSE
table_to_html(tbl, to_html = FALSE)
```

### 2023-04-15 posts/rtables/multiple_levels_calculation_2.qmd
Create multiple_levels_calculation_2.qmd

--- 2023-04-15 posts/rtables/multiple_levels_calculation_1.qmd
@@
 ---
-title: 'Mutiple level calculation 1'
-date: '2023-04-07'
+title: 'Mutiple level calculation 2'
+date: '2023-04-09'
 author: Qing Zou
 categories: ['R', 'rtables']
-description: 'Explore how to handle the mutiple calculation by rtables'
+description: 'Wrap the analysis function'
 execute:
@@
 
-## Multiple lelvels calculation
-
-`Multipile levels` calculation for categorical variable is the most common cases in Table, e.g. `PCSA`, `Exposure tables`.
-
-In this section, we will explore how to produce these output in `rtables`.
-
-Below is the test data:
-
 ```{r cal2}
-#| echo: true
+#| echo: false
 #| eval: true
@@
 ```
-```{r cal2sd }
-#| echo: false
-#| eval: true
-#| message: FALSE
-knitr::kable(df, "html")%>%
-  kableExtra::kable_styling(font_size = 16)
-```
-
-Now, we need to take the `PARAMCD` as 1st level, and the `CRIT` as 2nd level to calculate the frequency of each `TEST`, it will be a table like PCSA.
-
-# Method 1
-
-the `rtables::analyze()` will firstly be used in this case.
-
-To use this function, the data should be transformed from `long strcture(BDS)` to `wide`.
-
-Steps:
-
-1. take action on setting up for the `missing value` and the `factor`:
-
- - fill the missing value as `Missing_value_flag`, which will be subsequently handled in `rtables::analyze()`
- - use function `tidyr::pivot_wider` implement the transformation, the new columns are used as the analyzed variables `vars` of `rtables::analyze()`
- - `factor` the values by given level values, the levels is the format of output
 
 ```{r cal2df}
-#| echo: true
+#| echo: false
 #| eval: true
@@
 
-df1 <- df %>% tidyr::pivot_wider(id_cols = c("SUBJID", 'ASEQ', 'GROUP'),
-                                 names_from = "PARAMCD",values_from = "CRIT")
+df1 <- df %>% tidyr::pivot_wider(id_cols = c("SUBJID", 'ASEQ', 'GROUP'), names_from = "PARAMCD",values_from = "CRIT")
 
@@
 df1 <- df1 %>%
-  mutate(across(paste0('TEST_', 1:2), ~factor(fmt[[cur_column()]][.x],
-                                              levels = c(fmt[[cur_column()]])), .names = "{.col}_fmt"))
+  mutate(across(paste0('TEST_', 1:2), ~factor(fmt[[cur_column()]][.x], levels = c(fmt[[cur_column()]])), .names = "{.col}_fmt"))
+
+col_counts_df <- df1 %>%
+  distinct(SUBJID, GROUP)
+col_counts <- table(col_counts_df$GROUP)
+col_counts <- c(col_counts, sum(col_counts))
 ```
-```{r cal2g }
-#| echo: false
+
+## Wrap the analysis function
+
+The function is too long and it's not convient mixed with the layout of rtbales.
+For better user experience, we can create a function `summarise_vars_categorical`
+
+1. wrap the analysis function as `summarise_vars_categorical_afun`
+2. wrap the `summarise_vars_categorical_afun` and the `extral_args` within the `summarise_vars_categorical`
+
+Then user can passed all the analysis variables and customized argument by a single function `summarise_vars_categorical`.
+
+```{r a23s}
+#| echo: true
 #| eval: true
 #| message: FALSE
-knitr::kable(df1, "html")%>%
-  kableExtra::kable_styling(font_size = 16)
-```
 
-2. The dataframe `df1` is ready for output. now, the `layout` and `key values` of table will be built in the `rtables::analyze()`.
-
- - the header and overall column are generated by  `split_cols_by` and `add_overall_col`
- - in the `rtables::analyze()`:
-   - `vars` and `var_labels` are the variable names and labels
-   - in the `afun`, we create an analysis function, in which the arguments `df`, `.N_col`, `.var`, `.N_row` are created by rtables itself.
-the other aruments are customized will be passed by `extral_args`
-     - the `df` is current available(`non NA`) data with splited group, that's why we firstly fill the missing data `Missing_value_flag`. By filling
-       a non-NA value, we can handle the missing value as users expected.
-     - you will see the customized arguments `missing_rm` and `missing_label`, if user hope hide or remove the missing values.
-     - the `n_unique_keys` and `N_unique_keys` indicate that when counting the number of n/N, which records should be used.
-     - `add_Number` means whether add an row `Number: N`
-     - All the values of these customized arguments are passed by the `extra_args`, it's the greatest way to add the flexiabiliy for the tabulation.
-
-```{r cal2as}
-#| echo: true
-#| eval: true
-
-lyt <- basic_table() %>%
-  split_cols_by("GROUP") %>%
-  add_overall_col("All") %>%
-  rtables::analyze(vars = c('TEST_1_fmt', 'TEST_2_fmt'),
-                   var_labels = c('This is test 1', 'This is test 2'),
-                   afun = function(df, .N_col, .var, .N_row,
+# wrap the analysis function
+summarise_vars_categorical_afun <- function(df, .N_col, .var, .N_row,
                                             format = 'n(%)',
@@
                                             missing_rm = TRUE,
-                                  missing_label =NULL,
+                                            missing_label = NULL,
+                                            indent_mod = 0L,
                                             ...) {
@@
   }
-
   x <- df_n[[.var]]
-
   if(!is.null(N_unique_keys)){
@@
             format = sprintf_format_helper(xi, cus_N_col, 1, type = format),
-                              indent_mod = 1L
+            indent_mod = indent_mod
       )
@@
   )
+  if(add_Number){
+    lst_body <- append(setNames(list(rcell(cus_N_col,
+                                           format = "xx",
+                                           indent_mod = indent_mod)), 'Number'), lst_body)
+  }
 
@@
 
-                    if(add_Number){
-                      lst_body <- append(setNames(list(rcell(cus_N_col,
-                                                            format = "xx",
-                                                            indent_mod = 1L)),
-                                                  'Number'), lst_body)
-                    }
   in_rows(
@@
   )
-                  },
-          extra_args = list(format = 'n/N(%)',
+}
+
+# wrap the summarise_vars_categorical_afun and the extral_args
+summarise_vars_categorical <- function(lyt,
+                                       vars,
+                                       var_labels,
+                                       format = 'n(%)',
+                                       decimal = 1,
+                                       n_unique_keys = NULL,
+                                       N_unique_keys = NULL,
+                                       show_N_item = FALSE,
+                                       add_Number = FALSE,
+                                       missing_rm = TRUE,
+                                       missing_label = NULL,
+                                       show_labels='visible',
+                                       indent_mod = 0L, ...){
+  rtables::analyze(lyt, vars, var_labels,
+                   afun = summarise_vars_categorical_afun,
+                   extra_args = list(format = format,
+                                     decimal = decimal,
+                                     n_unique_keys = n_unique_keys,
+                                     N_unique_keys = N_unique_keys,
+                                     show_N_item = show_N_item,
+                                     add_Number = add_Number,
+                                     missing_rm = missing_rm,
+                                     missing_label = missing_label),
+                   show_labels = show_labels, indent_mod = indent_mod)
+}
+
+lyt <- basic_table() %>%
+  split_cols_by("GROUP") %>%
+  add_overall_col("All") %>%
+  summarise_vars_categorical(vars = c('TEST_1_fmt', 'TEST_2_fmt'),
+                             var_labels = c('This is test 1', 'This is test 2'),
+                             format = 'n/N(%)',
                              n_unique_keys = c("SUBJID"),
                              N_unique_keys = c("SUBJID"),
-                            add_Number = FALSE,
-                            missing_rm = FALSE,
-                            missing_label = NULL)
-          # ,
-          # show_labels = "hidden"
+                             add_Number = TRUE,
+                             missing_rm = FALSE
+
   )
-```
-3. finally, fit the `layout` and `df1` into the `rtables::build_table`
-
-```{r cal2ad}
-#| echo: true
-#| eval: true
-col_counts_df <- df1 %>%
-  distinct(SUBJID, GROUP)
-col_counts <- table(col_counts_df$GROUP)
-col_counts <- c(col_counts, sum(col_counts))
 
 tbl <- lyt %>% rtables::build_table(df1, col_counts = col_counts)
-```
-
-```{r cap1}
-#| echo: false
-#| eval: true
-my_caption <- ''
 ```

### 2023-04-15 posts/rtables/multiple_levels_calculation_3.qmd
Create multiple_levels_calculation_3.qmd

---
title: 'Mutiple level calculation 3'
date: '2023-04-10'
author: Qing Zou
categories: ['R', 'rtables']
description: 'summary counts by summarize_row_groups'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
format:
  html:
    page-layout: full

---

```{r rcell3sd45}
#| echo: false
#| eval: true

#' a pair of functions sprintf_format and format_values used for rcell
#'
#' @param n Unique number of events for the current category
#' @param N Total number of events under the split column
#' @param decimal Number of decimal to display
#' @param type to decide if the denominator n/N(%) in the result
#' @keywords sprintf_format
#'
#' @return
sprintf_format_helper <- function(n, N, decimal = 1, type = "n(%)") {
  showN <- NULL
  if (type == "n/N(%)"){
    showN <- "/%.0f"
  }

  if (n == 0) {
    fmt <- sprintf_format(paste0("%.0f", showN))
  } else if (n == N) {
    # if pct == 100%, then show (100) instead of (100.0)
    fmt <- sprintf_format(paste0("%.0f", showN, " ", "(%.0f)"))
  } else {
    fmt <- sprintf_format(paste0("%.0f", showN, " (%.", decimal, "f)"))
  }

  return(fmt)
}

#' a pair of functions sprintf_format and format_values used for rcell
#'
#' @param n Unique number of events for the current category
#' @param N Total number of events under the split column
#' @param type to decide if the denominator n/N(%) in the result
#'
#' @keywords format_values
#' @example
#' rcell(x = format_values_helper(1, 10),
#'       format = sprintf_format_helper(1, 10))
#'
#' rcell(x = format_values_helper(1, 10, type = "n/N(%)"),
#'       format = sprintf_format_helper(1, 10, type = "n/N(%)"))
#'
#' rcell(x = format_values_helper(10, 10),
#'       format = sprintf_format_helper(10, 10))
#'
#' @return
format_values_helper <- function(n, N, type = "n(%)") {
  showN <- NULL
  if (type == "n/N(%)"){
    showN <- N
  }
  if (n == 0) {
    value <- c(0, showN)
  } else {
    value <- c(n, showN, n / N * 100)
  }

  return(value)
}

#' Display rtables in html
#'
#' @param table a rtables object
#' @param to_html export to HTML. Default TRUE
#'
#' @return a html table
#' @export
table_to_html <- function(table, to_html = TRUE) {
  stopifnot(!is.null(table), class(table) %in% c("TableTree", "ElementaryTable"))

  extra_css <- "
  .tabwid {overflow-x: scroll; display: block;}
  .tabwid table td p {min-width: 170px !important;}
  .tabwid table td:first-child > p {min-width: 200px !important;}
  .tabwid table tr:hover {background-color: #ffff99 !important;}
  "
  flextable::set_flextable_defaults(
    font.family = ifelse(to_html, "Source Sans Pro", "Verdana"),
    extra_css = extra_css
  )

  ft <- rtables::tt_to_flextable(table) %>%
    flextable::border_remove() %>%
    flextable::bold(bold = TRUE, part = "header") %>%
    flextable::hline_top(part = "body") %>%
    flextable::hline_bottom(part = "body") %>%
    flextable::autofit()

  if (to_html) {
    flextable::htmltools_value(ft)
  } else {
    ft
  }
}
count_items_cfun <- function(df,
                        labelstr,
                        .spl_context,
                        n_var,
                        n_unique_keys = NULL,
                        N_var = NULL,
                        N_index = NULL,
                        N_unique_keys = NULL,
                        empty_row = FALSE,
                        hide_value = FALSE,
                        format = "n/N(%)") {

  # caculate the value
  df_n <- df
  if (!is.null(n_unique_keys)) {
    df_n <- df %>% distinct(across(n_unique_keys), .keep_all = T)
  }
  x <- df_n[[n_var]]
  n <- sum(x[!is.na(x)])

  cur_index <- which(.spl_context[["value"]] == labelstr)
  # View(df)
  if (!is.null(N_var)) {

    # get N of parent for denomitor caculation
    if (!is.null(N_index)) {
      parent_df <- .spl_context$full_parent_df[[N_index]][.spl_context[["cur_col_subset"]][[N_index]], ]
    } else {
      parent_df <- .spl_context$full_parent_df[[cur_index - 1]][.spl_context[["cur_col_subset"]][[cur_index - 1]], ]
    }

    if (!is.null(N_unique_keys)) {
      parent_df <-  parent_df %>% distinct(across(N_unique_keys), .keep_all = T)
    }

    parent_x <- parent_df[[N_var]]
    parent_N <- sum(!is.na(parent_x[!is.na(parent_x)]))

    # get current value to check if it match the last value of levels, for empty row
    # !Start
    var_levels <- length(.spl_context$split)
    curr_var_level_val <- .spl_context$value[[nrow(.spl_context)]]
    last_var_level_val <-
      .spl_context$full_parent_df[[nrow(.spl_context)]][[.spl_context$split[var_levels]]] %>%
      levels() %>%
      tail(1)
    if (curr_var_level_val == last_var_level_val & empty_row) {
      ret <- in_rows(
        .names = c(labelstr, " "),
        .list = list(c(format_values_helper(n, parent_N, type = format)), NULL),
        .formats = c(sprintf_format_helper(n, parent_N, type = format), "xx")
      )
    } else {
      rc <- rcell(
        x = format_values_helper(n, parent_N, type = format),
        format = sprintf_format_helper(n, parent_N, type = format)
      )
      ret <- in_rows(.names = labelstr, rc)
    }
    # !End
  } else {
    rc <- rcell(
      x = ifelse(hide_value, "", n),
      format = "xx"
    )
    ret <- in_rows(.names = labelstr, rc)
  }
  return(ret)
}
```

```{r cal1}
library(rtables)
library(formatters)
library(dplyr)
library(stringr)
```

## Method 2

```{r cal2}
#| echo: false
#| eval: true
df <- data.frame(SUBJID = c(rep('S1', 5), rep("S2", 4), rep("S3", 2)),
                 ASEQ = c(1:5, 1:4, 1:2),
                 GROUP = factor(c(rep('A', 5), rep("B", 4), rep('A', 2)), levels = c("A", "B")),
                 PARAMCD = c('TEST_1', 'TEST_1', 'TEST_1', 'TEST_1', 'TEST_2', 'TEST_1', 'TEST_1', 'TEST_2', 'TEST_2', 'TEST_1', 'TEST_2'),
                 CRIT = c('LEVEL_2_1','LEVEL_2_1', 'LEVEL_2_2', NA_character_, 'LEVEL_2_a', 'LEVEL_2_1', NA_character_, 'LEVEL_2_a', 'LEVEL_2_b', NA_character_, 'LEVEL_2_b'))
```
Another powerful function rtables provide to us is the function `summarize_row_groups`, which can be used to add a content row of summary counts.

This function need a different data structure from the `rtable::analyze()`.
`rtable::analyze()` is kind of the columns-based analysis, the `summarize_row_groups` is more like the rows-based analysis.

To build the data:

- create the 1st level columns by `PARAMCD`, `TEST_1_level_1`, `TEST_2_level_1`
- Create the 2rd level columns by `CRIT`, `TEST_1_level_2`, `TEST_2_level_2`

```{r cal2df}
#| echo: true
#| eval: true

param_coding <- c('TEST_1' = 'This is test 1',
                  'TEST_2' = 'This is test 2')
df <- df %>% mutate(PARAML = param_coding[PARAMCD])
tmp <- df
tmp[, str_c(names(param_coding), '_level_1')] <- NA_character_

# tmp1 <- df %>% mutate(TEST_1_level_1 = ifelse(PARAMCD == 'TEST_1', PARAMCD, NA_character_),
#                       TEST_2_level_1 = ifelse(PARAMCD == 'TEST_2', PARAMCD, NA_character_),
#                       TEST_1_aval_1 = ifelse(!is.na(TEST_1_level_1), 1, NA_character_),
#                       TEST_1_aval_1 = ifelse(!is.na(TEST_2_level_1), 1, NA_character_),
#                       TEST_1_level_2 = ifelse(PARAMCD == 'TEST_1' & !is.na(CRIT), CRIT, NA_character_),
#                       TEST_2_level_2 = ifelse(PARAMCD == 'TEST_2' & !is.na(CRIT), CRIT, NA_character_),
#                       TEST_1_aval_2 = ifelse(!is.na(TEST_1_level_2), 1, NA_character_),
#                       TEST_2_aval_2 = ifelse(!is.na(TEST_2_level_2), 1, NA_character_)
#                       )

tmp1 <- tmp %>%
  mutate(
    across(
      .cols = c("TEST_1_level_1", "TEST_2_level_1"),
      .fns = \(x)  factor(ifelse(grepl(gsub("_level_1", "",cur_column()), PARAMCD), PARAML, x))
    )
    ,
    across(c(do.call(str_c, expand.grid(names(param_coding), c('_level_1')))),
           ~ifelse(!is.na(.x), 1, .x),
           .names = "{gsub('level','aval', .col)}")
    )

tmp1[, str_c(names(param_coding), '_level_2')] <- NA_character_

fmt <- list(
  'TEST_1_level_2' = c('LEVEL_2_1' = '1 of level 2',
               'LEVEL_2_2' = '2 of level 2',
               'LEVEL_2_3' = '3 of level 2'),
  "TEST_2_level_2" = c('LEVEL_2_a' = 'a of level 2',
               'LEVEL_2_b' = 'b of level 2',
               'LEVEL_2_c' = 'c of level 2')
)

tmp2 <- tmp1 %>%
  mutate(
    across(
      .cols = c("TEST_1_level_2", "TEST_2_level_2"),
      .fns = \(x)  factor(ifelse(grepl(gsub("_level_2", "",cur_column()), PARAMCD) & !is.na(CRIT) , fmt[[cur_column()]][CRIT], x), levels = fmt[[cur_column()]])
    ),
    across(c(do.call(str_c, expand.grid(names(param_coding), c('_level_2')))),
           ~ifelse(!is.na(.x), 1, .x),
           .names = "{gsub('level','aval', .col)}")
  )
```

## Wrap the analysis function

```{r a23s}
#| echo: true
#| eval: true
#| message: FALSE

tbl_recipe <- basic_table() %>%
  rtables::split_cols_by(var = "GROUP") %>%
  add_overall_col("All") %>%
  ## 1st-level row-wise splitting, and its column-wise analysis function
  rtables::split_rows_by("TEST_1_level_1", split_fun = remove_split_levels(c(NA_character_))) %>%
  rtables::summarize_row_groups(cfun = count_items_cfun,
                                extra_args = list(n_var = "TEST_1_aval_1", n_unique_keys = "SUBJID", hide_value = TRUE)) %>%
  ## 2nd-level row-wise splitting, and its column-wise analysis function, use n_unique_keys to remove duplicated records
  rtables::split_rows_by("TEST_1_level_2") %>%
  rtables::summarize_row_groups(cfun = count_items_cfun, extra_args = list(
    n_var = "TEST_1_aval_2", n_unique_keys = c("SUBJID"), N_var = "TEST_1_aval_1", N_unique_keys = "SUBJID",
    empty_row = TRUE
  ))

col_counts_df <- tmp2 %>%
  distinct(SUBJID, GROUP)
col_counts <- table(col_counts_df$GROUP)
col_counts <- c(col_counts, sum(col_counts))
test1_tbl <- tbl_recipe %>%
  rtables::build_table(tmp2, col_counts = col_counts)
```
```{r}
#| echo: false
#| eval: true
#| message: FALSE
table_to_html(test1_tbl, to_html = FALSE)
```

```{r a23ssdf}
#| echo: true
#| eval: true
#| message: FALSE

count_items_cfun <- function(df,
                        labelstr,
                        .spl_context,
                        n_var,
                        n_unique_keys = NULL,
                        N_var = NULL,
                        N_index = NULL,
                        N_unique_keys = NULL,
                        empty_row = FALSE,
                        hide_value = FALSE,
                        format = "n/N(%)") {

  # caculate the value
  df_n <- df
  if (!is.null(n_unique_keys)) {
    df_n <- df %>% distinct(across(n_unique_keys), .keep_all = T)
  }
  x <- df_n[[n_var]]
  n <- sum(x[!is.na(x)])

  cur_index <- which(.spl_context[["value"]] == labelstr)
  # View(df)
  if (!is.null(N_var)) {

    # get N of parent for denomitor caculation
    if (!is.null(N_index)) {
      parent_df <- .spl_context$full_parent_df[[N_index]][.spl_context[["cur_col_subset"]][[N_index]], ]
    } else {
      parent_df <- .spl_context$full_parent_df[[cur_index - 1]][.spl_context[["cur_col_subset"]][[cur_index - 1]], ]
    }

    if (!is.null(N_unique_keys)) {
      parent_df <-  parent_df %>% distinct(across(N_unique_keys), .keep_all = T)
    }

    parent_x <- parent_df[[N_var]]
    parent_N <- sum(!is.na(parent_x[!is.na(parent_x)]))

    # get current value to check if it match the last value of levels, for empty row
    # !Start
    var_levels <- length(.spl_context$split)
    curr_var_level_val <- .spl_context$value[[nrow(.spl_context)]]
    last_var_level_val <-
      .spl_context$full_parent_df[[nrow(.spl_context)]][[.spl_context$split[var_levels]]] %>%
      levels() %>%
      tail(1)
    if (curr_var_level_val == last_var_level_val & empty_row) {
      ret <- in_rows(
        .names = c(labelstr, " "),
        .list = list(c(format_values_helper(n, parent_N, type = format)), NULL),
        .formats = c(sprintf_format_helper(n, parent_N, type = format), "xx")
      )
    } else {
      rc <- rcell(
        x = format_values_helper(n, parent_N, type = format),
        format = sprintf_format_helper(n, parent_N, type = format)
      )
      ret <- in_rows(.names = labelstr, rc)
    }
    # !End
  } else {
    rc <- rcell(
      x = ifelse(hide_value, "", n),
      format = "xx"
    )
    ret <- in_rows(.names = labelstr, rc)
  }
  return(ret)
}

tbl_recipe <- basic_table() %>%
  rtables::split_cols_by(var = "GROUP") %>%
  add_overall_col("All") %>%
  ## 1st-level row-wise splitting, and its column-wise analysis function
  rtables::split_rows_by("TEST_2_level_1", split_fun = remove_split_levels(c(NA_character_))) %>%
  rtables::summarize_row_groups(cfun = count_items_cfun,
                                extra_args = list(n_var = "TEST_2_aval_1", n_unique_keys = "SUBJID", hide_value = TRUE)) %>%
  ## 2nd-level row-wise splitting, and its column-wise analysis function, use n_unique_keys to remove duplicated records
  rtables::split_rows_by("TEST_2_level_2") %>%
  rtables::summarize_row_groups(cfun = count_items_cfun, extra_args = list(
    n_var = "TEST_2_aval_2", n_unique_keys = c("SUBJID"), N_var = "TEST_2_aval_1", N_unique_keys = "SUBJID",
    empty_row = TRUE
  ))

test2_tbl <- tbl_recipe %>%
  rtables::build_table(tmp2, col_counts = col_counts)
```
```{r}
#| echo: false
#| eval: true
#| message: FALSE
table_to_html(test2_tbl, to_html = FALSE)
```

```{r a23ssdfdf}
#| echo: true
#| eval: true
#| message: FALSE

tbl2 <- rbind(test1_tbl, test2_tbl)

```

```{r}
#| echo: false
#| eval: true
#| message: FALSE
table_to_html(tbl2, to_html = FALSE)
```

### 2023-04-15 posts/rtables/multiple_levels_calculation_4.qmd
Create multiple_levels_calculation_4.qmd

--- 2023-04-15 posts/rtables/multiple_levels_calculation_3.qmd
@@
 ---
-title: 'Mutiple level calculation 3'
-date: '2023-04-10'
+title: 'Mutiple level calculation 4'
+date: '2023-04-11'
 author: Qing Zou
 categories: ['R', 'rtables']
-description: 'summary counts by summarize_row_groups'
+description: 'multiple level counts more than 2 levels'
 execute:
@@
   chunk_output_type: console
-format:
-  html:
-    page-layout: full
-
 ---
@@
 }
+
 count_items_cfun <- function(df,
@@
 
-## Method 2
+## Summary counts more than 2rd level
 
@@
 #| eval: true
-df <- data.frame(SUBJID = c(rep('S1', 5), rep("S2", 4), rep("S3", 2)),
-                 ASEQ = c(1:5, 1:4, 1:2),
-                 GROUP = factor(c(rep('A', 5), rep("B", 4), rep('A', 2)), levels = c("A", "B")),
-                 PARAMCD = c('TEST_1', 'TEST_1', 'TEST_1', 'TEST_1', 'TEST_2', 'TEST_1', 'TEST_1', 'TEST_2', 'TEST_2', 'TEST_1', 'TEST_2'),
-                 CRIT = c('LEVEL_2_1','LEVEL_2_1', 'LEVEL_2_2', NA_character_, 'LEVEL_2_a', 'LEVEL_2_1', NA_character_, 'LEVEL_2_a', 'LEVEL_2_b', NA_character_, 'LEVEL_2_b'))
+df <- data.frame(SUBJID = paste0("S", 1:10),
+                 GROUP = factor(sample(c("A", "B", "C"), 10, replace = T), levels = c("A", "B", "C")),
+                 AVAL = sample.int(20, 10)
+                 )
+
+df <- df %>% mutate(level_1 = ifelse(3 < AVAL & AVAL <=18, "level 1", NA_character_),
+                    level_2 = ifelse(5 < AVAL & AVAL <=15, "level 2", NA_character_),
+                    level_3 = ifelse(8 < AVAL & AVAL  <=12, "level_3", NA_character_)
+                    )
+df <- df %>% mutate(across(paste0('level_', 1:3),
+                           ~factor(.)))
+df <- df %>% mutate(across(paste0('level_', 1:3),
+                           ~ifelse(!is.na(.x), 1, .x),
+                           .names = "{gsub('level','aval', .col)}"))
 ```
-Another powerful function rtables provide to us is the function `summarize_row_groups`, which can be used to add a content row of summary counts.
+```{r cal2sd }
+#| echo: false
+#| eval: true
+#| message: FALSE
+knitr::kable(df, "html")%>%
+  kableExtra::kable_styling(font_size = 16)
+```
 
-This function need a different data structure from the `rtable::analyze()`.
-`rtable::analyze()` is kind of the columns-based analysis, the `summarize_row_groups` is more like the rows-based analysis.
-
-To build the data:
-
-- create the 1st level columns by `PARAMCD`, `TEST_1_level_1`, `TEST_2_level_1`
-- Create the 2rd level columns by `CRIT`, `TEST_1_level_2`, `TEST_2_level_2`
-
-```{r cal2df}
+```{r cal2dfg}
 #| echo: true
@@
 
-param_coding <- c('TEST_1' = 'This is test 1',
-                  'TEST_2' = 'This is test 2')
-df <- df %>% mutate(PARAML = param_coding[PARAMCD])
-tmp <- df
-tmp[, str_c(names(param_coding), '_level_1')] <- NA_character_
-
-# tmp1 <- df %>% mutate(TEST_1_level_1 = ifelse(PARAMCD == 'TEST_1', PARAMCD, NA_character_),
-#                       TEST_2_level_1 = ifelse(PARAMCD == 'TEST_2', PARAMCD, NA_character_),
-#                       TEST_1_aval_1 = ifelse(!is.na(TEST_1_level_1), 1, NA_character_),
-#                       TEST_1_aval_1 = ifelse(!is.na(TEST_2_level_1), 1, NA_character_),
-#                       TEST_1_level_2 = ifelse(PARAMCD == 'TEST_1' & !is.na(CRIT), CRIT, NA_character_),
-#                       TEST_2_level_2 = ifelse(PARAMCD == 'TEST_2' & !is.na(CRIT), CRIT, NA_character_),
-#                       TEST_1_aval_2 = ifelse(!is.na(TEST_1_level_2), 1, NA_character_),
-#                       TEST_2_aval_2 = ifelse(!is.na(TEST_2_level_2), 1, NA_character_)
-#                       )
-
-tmp1 <- tmp %>%
-  mutate(
-    across(
-      .cols = c("TEST_1_level_1", "TEST_2_level_1"),
-      .fns = \(x)  factor(ifelse(grepl(gsub("_level_1", "",cur_column()), PARAMCD), PARAML, x))
-    )
-    ,
-    across(c(do.call(str_c, expand.grid(names(param_coding), c('_level_1')))),
-           ~ifelse(!is.na(.x), 1, .x),
-           .names = "{gsub('level','aval', .col)}")
-    )
-
-tmp1[, str_c(names(param_coding), '_level_2')] <- NA_character_
-
-fmt <- list(
-  'TEST_1_level_2' = c('LEVEL_2_1' = '1 of level 2',
-               'LEVEL_2_2' = '2 of level 2',
-               'LEVEL_2_3' = '3 of level 2'),
-  "TEST_2_level_2" = c('LEVEL_2_a' = 'a of level 2',
-               'LEVEL_2_b' = 'b of level 2',
-               'LEVEL_2_c' = 'c of level 2')
-)
-
-tmp2 <- tmp1 %>%
-  mutate(
-    across(
-      .cols = c("TEST_1_level_2", "TEST_2_level_2"),
-      .fns = \(x)  factor(ifelse(grepl(gsub("_level_2", "",cur_column()), PARAMCD) & !is.na(CRIT) , fmt[[cur_column()]][CRIT], x), levels = fmt[[cur_column()]])
-    ),
-    across(c(do.call(str_c, expand.grid(names(param_coding), c('_level_2')))),
-           ~ifelse(!is.na(.x), 1, .x),
-           .names = "{gsub('level','aval', .col)}")
-  )
-```
-
-## Wrap the analysis function
-
-```{r a23s}
-#| echo: true
-#| eval: true
-#| message: FALSE
+col_counts <- table(df$GROUP)
+col_counts <- c(col_counts, sum(col_counts))
 
@@
   ## 1st-level row-wise splitting, and its column-wise analysis function
-  rtables::split_rows_by("TEST_1_level_1", split_fun = remove_split_levels(c(NA_character_))) %>%
+  rtables::split_rows_by("level_1", split_fun = remove_split_levels(c(NA_character_))) %>%
   rtables::summarize_row_groups(cfun = count_items_cfun,
-                                extra_args = list(n_var = "TEST_1_aval_1", n_unique_keys = "SUBJID", hide_value = TRUE)) %>%
+                                extra_args = list(n_var = "aval_1", n_unique_keys = "SUBJID", hide_value = FALSE)) %>%
   ## 2nd-level row-wise splitting, and its column-wise analysis function, use n_unique_keys to remove duplicated records
-  rtables::split_rows_by("TEST_1_level_2") %>%
+  rtables::split_rows_by("level_2") %>%
   rtables::summarize_row_groups(cfun = count_items_cfun, extra_args = list(
-    n_var = "TEST_1_aval_2", n_unique_keys = c("SUBJID"), N_var = "TEST_1_aval_1", N_unique_keys = "SUBJID",
+    n_var = "aval_2", N_var = "aval_1", N_unique_keys = "SUBJID",
+    n_unique_keys = c("SUBJID"), empty_row = FALSE
+  )) %>%
+  ## 3rd-level row-wise splitting, and its column-wise analysis function, use n_unique_keys to remove duplicated records
+  rtables::split_rows_by("level_3") %>%
+  rtables::summarize_row_groups(cfun = count_items_cfun, extra_args = list(
+    n_var = "aval_3", n_unique_keys = c("SUBJID"), N_var = "aval_1", N_unique_keys = "SUBJID", N_index = 1,
     empty_row = TRUE
@@
 
-col_counts_df <- tmp2 %>%
-  distinct(SUBJID, GROUP)
-col_counts <- table(col_counts_df$GROUP)
-col_counts <- c(col_counts, sum(col_counts))
-test1_tbl <- tbl_recipe %>%
-  rtables::build_table(tmp2, col_counts = col_counts)
-```
-```{r}
-#| echo: false
-#| eval: true
-#| message: FALSE
-table_to_html(test1_tbl, to_html = FALSE)
-```
-
-```{r a23ssdf}
-#| echo: true
-#| eval: true
-#| message: FALSE
-
-count_items_cfun <- function(df,
-                        labelstr,
-                        .spl_context,
-                        n_var,
-                        n_unique_keys = NULL,
-                        N_var = NULL,
-                        N_index = NULL,
-                        N_unique_keys = NULL,
-                        empty_row = FALSE,
-                        hide_value = FALSE,
-                        format = "n/N(%)") {
-
-  # caculate the value
-  df_n <- df
-  if (!is.null(n_unique_keys)) {
-    df_n <- df %>% distinct(across(n_unique_keys), .keep_all = T)
-  }
-  x <- df_n[[n_var]]
-  n <- sum(x[!is.na(x)])
-
-  cur_index <- which(.spl_context[["value"]] == labelstr)
-  # View(df)
-  if (!is.null(N_var)) {
-
-    # get N of parent for denomitor caculation
-    if (!is.null(N_index)) {
-      parent_df <- .spl_context$full_parent_df[[N_index]][.spl_context[["cur_col_subset"]][[N_index]], ]
-    } else {
-      parent_df <- .spl_context$full_parent_df[[cur_index - 1]][.spl_context[["cur_col_subset"]][[cur_index - 1]], ]
-    }
-
-    if (!is.null(N_unique_keys)) {
-      parent_df <-  parent_df %>% distinct(across(N_unique_keys), .keep_all = T)
-    }
-
-    parent_x <- parent_df[[N_var]]
-    parent_N <- sum(!is.na(parent_x[!is.na(parent_x)]))
-
-    # get current value to check if it match the last value of levels, for empty row
-    # !Start
-    var_levels <- length(.spl_context$split)
-    curr_var_level_val <- .spl_context$value[[nrow(.spl_context)]]
-    last_var_level_val <-
-      .spl_context$full_parent_df[[nrow(.spl_context)]][[.spl_context$split[var_levels]]] %>%
-      levels() %>%
-      tail(1)
-    if (curr_var_level_val == last_var_level_val & empty_row) {
-      ret <- in_rows(
-        .names = c(labelstr, " "),
-        .list = list(c(format_values_helper(n, parent_N, type = format)), NULL),
-        .formats = c(sprintf_format_helper(n, parent_N, type = format), "xx")
-      )
-    } else {
-      rc <- rcell(
-        x = format_values_helper(n, parent_N, type = format),
-        format = sprintf_format_helper(n, parent_N, type = format)
-      )
-      ret <- in_rows(.names = labelstr, rc)
-    }
-    # !End
-  } else {
-    rc <- rcell(
-      x = ifelse(hide_value, "", n),
-      format = "xx"
-    )
-    ret <- in_rows(.names = labelstr, rc)
-  }
-  return(ret)
-}
-
-tbl_recipe <- basic_table() %>%
-  rtables::split_cols_by(var = "GROUP") %>%
-  add_overall_col("All") %>%
-  ## 1st-level row-wise splitting, and its column-wise analysis function
-  rtables::split_rows_by("TEST_2_level_1", split_fun = remove_split_levels(c(NA_character_))) %>%
-  rtables::summarize_row_groups(cfun = count_items_cfun,
-                                extra_args = list(n_var = "TEST_2_aval_1", n_unique_keys = "SUBJID", hide_value = TRUE)) %>%
-  ## 2nd-level row-wise splitting, and its column-wise analysis function, use n_unique_keys to remove duplicated records
-  rtables::split_rows_by("TEST_2_level_2") %>%
-  rtables::summarize_row_groups(cfun = count_items_cfun, extra_args = list(
-    n_var = "TEST_2_aval_2", n_unique_keys = c("SUBJID"), N_var = "TEST_2_aval_1", N_unique_keys = "SUBJID",
-    empty_row = TRUE
-  ))
-
-test2_tbl <- tbl_recipe %>%
-  rtables::build_table(tmp2, col_counts = col_counts)
-```
-```{r}
-#| echo: false
-#| eval: true
-#| message: FALSE
-table_to_html(test2_tbl, to_html = FALSE)
-```
-
-```{r a23ssdfdf}
-#| echo: true
-#| eval: true
-#| message: FALSE
-
-tbl2 <- rbind(test1_tbl, test2_tbl)
-
+tbl_test3 <- tbl_recipe %>%
+  rtables::build_table(df, col_counts = col_counts)
 ```
@@
 #| message: FALSE
-table_to_html(tbl2, to_html = FALSE)
+table_to_html(tbl_test3, to_html = FALSE)
 ```

### 2023-04-15 posts/manipulate/dplyr_cross.qmd
Create dplyr_cross.qmd

---
title: 'Derive multiple columns based on across'
date: '2023-04-08'
author: Qing Zou
categories: ['R', 'dplyr', 'across']
description: 'When we want to add new columns and have to iterate...'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

```{r rcell0}
library(dplyr)
```

## Reale Case

Suppose we this data with columns `num_var1_1`, `num_var1_99` and `denom1_1`, `denom1_99`

```{r rcell15}
#| echo: true
#| eval: true
test_data <- data.frame(num_var1_1 = c(1, 4),
                        num_var1_99 = c(2, 6),
                        denom1_1 = c(2, 5),
                        denom1_99 = c(3, 1))
```

what we want to do is to add two new columns: the new column equal to the sum of `num_var_x` and `denom1_y`.
Just fit the data into the function `sum`

```{r rcell116}
#| echo: true
#| eval: true

sum_c <- function(x, y){
  ifelse(y > 2, x+y+10, x+y)
}

test_data1 <- test_data %>%
             mutate(var1_1 = sum_c(num_var1_1, denom1_1),
                    var1_99 = sum_c(num_var1_99, denom1_99))
```

```{r rcell11612 }
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(test_data1, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

the code seems not readable.

## dplyr::across()

we can use the across() to easily apply the functions to the selected columns.

__Description__

`across()` makes it easy to apply the same transformation to multiple columns, allowing you to use `select()` semantics inside in "data-masking" functions like `summarise()` and `mutate()`. See `vignette("colwise")` for more details.

__Usage__

```{r rcell13}
#| echo: true
#| eval: false
across(.cols = everything(), .fns = NULL, ..., .names = NULL)
```

__Arguments__

  |
 -|:--
 __.cols__| `<tidy-select>` Columns to transform. You can't select grouping columns because they are already automatically handled by the verb (i.e. `summarise()` or `mutate()`).
 __.fns__| Functions to apply to each of the selected columns. Possible values are:

- A function, e.g. `mean`.

- A purrr-style lambda, e.g. `~ mean(.x, na.rm = TRUE)`

- A list of functions/lambdas, e.g. `list(mean = mean, n_miss = ~ sum(is.na(.x))`

- `NULL`: the default value, returns the selected columns in a data frame without applying a transformation. This is useful for when you want to use a function that takes a data frame.

Within these functions you can use `cur_column()` and `cur_group()` to access the current column and grouping keys respectively.
 __.names__|  A glue specification that describes how to name the output columns. This can use `{.col}` to stand for the selected column name, and `{.fn}` to stand for the name of the function being applied. The default `NULL` is equivalent to `{.col}` for the single function case and `{.col}_{.fn}` for the case where a list is used for .fns.

### Method 1

To use the `across` in this case:

```{r rcell2}
#| echo: true
#| eval: false
test_data %>%
  mutate(across(.cols = starts_with('num_var1'),
                .fns = ~sum_c(., get(gsub('num_var1', "denom1", cur_column()))),
                .names = "{gsub('num_var1', 'var1', .col)}"))
```
```{r rcell11612sd }
#| echo: false
#| eval: true
#| message: FALSE
test_data %>%
  mutate(across(.cols = starts_with('num_var1'),
                .fns = ~sum_c(., get(gsub('num_var1', "denom1", cur_column()))),
                .names = "{gsub('num_var1', 'var1', .col)}")) %>%
knitr::kable("html")%>%
  kableExtra::kable_styling(font_size = 16)
```

Steps:

1. select referenced columns by `starts_with('num_var1')`
2. apply the function `sum_c`. here, the first argument `x` is from the `.cols`, the second argument `y` is retrived by using the function
`cur_column()`.
3. give the names to new columns

### Method 2

this code will be used for a function that can do the vectorized calculation. it's faster that the method 1.

```{r rcell21}
#| echo: true
#| eval: false
test_data %>%
  mutate(Vectorize(sum_c)(across(starts_with('num_var1'),
                                 .names = "{gsub('num_var1', 'var1', .col)}"),
                          across(starts_with('denom1')))) %>%
  do.call(data.frame, .) %>%
  rename_at(dplyr::vars(starts_with("Vectorize")), ~ gsub("Vectorize.sum_c.......", "", .))
```
```{r rcell116112sd }
#| echo: false
#| eval: true
#| message: FALSE
test_data %>%
  mutate(Vectorize(sum_c)(across(starts_with('num_var1'),
                                 .names = "{gsub('num_var1', 'var1', .col)}"),
                          across(starts_with('denom1')))) %>%
  do.call(data.frame, .) %>%
  rename_at(dplyr::vars(starts_with("Vectorize")), ~ gsub("Vectorize.sum_c.......", "", .)) %>%
knitr::kable("html")%>%
  kableExtra::kable_styling(font_size = 16)
```

Steps:

1. in the `mutate`, the function firstly is vectorized since we need a vector result returned.
2. the function `sum_c` take the first `across()` as argument `x`, the name of new columns is set in here. the second `across()`
select the needed columns as argument `y`.
3. both passed value `x`, `y`  actually are two `dataframe`. e.g. the first one is `test_data %>% select(starts_with('num_var1'))`, the second one is `test_data %>% select(starts_with('denom1'))`. to accept the `dataframe` value, the  `sum_c` need to be a vectorized function.
4. the `mutate` return a `matrix` as the new column, which can be transformed by `data.frame`(`as.data.frame` not work)

### 2023-04-15 R/control.R
add control.R

@@ -1,0 +1,14 @@
+index_file <- "_site/index.html"
+pagecryptr::pagecryptr(file, "r", out_file = "_site/index.html")
+contents <- paste(readLines("_site/index.html"), collapse = "\n")
+writeLines(gsub('<div id=\"attribution\">\n        Protected by <a href=\"https://www.maxlaumeister.com/pagecrypt/\">PageCrypt</a>\n    </div>', "", contents), "_site/index.html")
+for (i in list.dirs('_site/posts')[-1]) {
+    posts <- list.files(i)
+    for (j in posts) {
+        file <- file.path(i, j)
+        pagecryptr::pagecryptr(file, "r", out_file = file)
+        contents <- paste(readLines(file), collapse = "\n")
+        writeLines(gsub('<div id=\"attribution\">\n        Protected by <a href=\"https://www.maxlaumeister.com/pagecrypt/\">PageCrypt</a>\n    </div>', "", contents), file)
+    }
+    
+}

### 2023-04-15 R/control.R
Update control.R

@@ -1,8 +1,8 @@
-index_file <- "_site/index.html"
-pagecryptr::pagecryptr(file, "r", out_file = "_site/index.html")
-contents <- paste(readLines("_site/index.html"), collapse = "\n")
-writeLines(gsub('<div id=\"attribution\">\n        Protected by <a href=\"https://www.maxlaumeister.com/pagecrypt/\">PageCrypt</a>\n    </div>', "", contents), "_site/index.html")
-for (i in list.dirs('_site/posts')[-1]) {
+index_file <- "index.html"
+pagecryptr::pagecryptr(file, "r", out_file = "index.html")
+contents <- paste(readLines("index.html"), collapse = "\n")
+writeLines(gsub('<div id=\"attribution\">\n        Protected by <a href=\"https://www.maxlaumeister.com/pagecrypt/\">PageCrypt</a>\n    </div>', "", contents), "index.html")
+for (i in list.dirs('posts')[-1]) {
     posts <- list.files(i)
     for (j in posts) {
         file <- file.path(i, j)

### 2023-04-15 R/control.R
Update control.R

@@ -1,5 +1,5 @@
-index_file <- "index.html"
-pagecryptr::pagecryptr(file, "r", out_file = "index.html")
+# Encrypt an HTML Page by pagecryptr
+pagecryptr::pagecryptr("index.html", "r", out_file = "index.html")
 contents <- paste(readLines("index.html"), collapse = "\n")
 writeLines(gsub('<div id=\"attribution\">\n        Protected by <a href=\"https://www.maxlaumeister.com/pagecrypt/\">PageCrypt</a>\n    </div>', "", contents), "index.html")
 for (i in list.dirs('posts')[-1]) {

### 2023-04-16 posts/rtables/customize_cell.qmd
fix typo

@@
 colspan = 1L,
-rcell 103
 label = NULL,

### 2023-04-16 posts/manipulate/formula_derive.R
Create formula_derive.R

@@ -1,0 +1,194 @@
+---
+title: 'Derive new columns based on given formula'
+date: '2023-04-16'
+author: Qing Zou
+categories: ['R', 'dplyr', 'glue', 'Derive']
+description: 'Efficiently Derive new columns based on formula'
+execute:
+  message: false
+  warning: false
+editor_options:
+  chunk_output_type: console
+---
+
+
+```{r rcell0}
+library(dplyr)
+library(glue)
+```
+
+## Real Case
+
+It is the most common in the data manipulation that deriving a column based on a given formula, such as `Toxicity Grade variables`, and `PCSA variables`.
+
+```{r rcell15}
+#| echo: true
+#| eval: true
+df <- data.frame(SUBJID = paste0("S", 1:10),
+                 GROUP = factor(sample(c("A", "B", "C"), 10, replace = T), levels = c("A", "B", "C")),
+                 AVAL = sample.int(20, 10)
+)
+```
+
+If we need to create new variables based on AVAL variable and condition or criteria, we have to do the following derivation.
+
+```{r}
+#| echo: true
+#| eval: true
+df <- df %>% mutate(level_1 = ifelse(AVAL>3 & GROUP == "A", "level 1", NA_character_),
+                    level_2 = ifelse(AVAL<3, "level 2", NA_character_))
+```
+```{r rcell11612 }
+#| echo: false
+#| eval: true
+#| message: FALSE
+knitr::kable(df, "html")%>%
+  kableExtra::kable_styling(font_size = 16)
+```
+
+we used the `ifelse` or `case_when` to implemente.In real case, the massive new variables and conditions lead to a not easy life!
+here, we can develop functions for functional programmer purpose to handle this kind of task.
+
+## Method 1
+
+we store the column names, formula, and value in a list `derived_rule`.
+```{r rcell116}
+#| echo: true
+#| eval: true
+
+derived_rule <- list('level_1_new' = list('formula' = 'AVAL>3 & GROUP == "A"', value = "level 1"),
+                'level_2_new' = list('formula' = 'AVAL<3', value = "level 2"))
+```
+
+we can combine the `glue` to build a function based on the given formula, and then make use of the `do.call` acheive our goal.
+
+##### 1. convert formula to a function
+
+```{r}
+#| echo: true
+#| eval: true
+convert_to_function <- function(formula, value = NULL) {
+  args <- paste(all.vars(rlang::parse_expr(formula)), collapse = ", ")
[…]

### 2023-04-16 posts/manipulate/formula_derive.qmd
Update formula_derive.qmd

---
title: 'Derive new columns based on given formula'
date: '2023-04-16'
author: Qing Zou
categories: ['R', 'dplyr', 'glue', 'Derive']
description: 'Efficiently Derive new columns based on formula'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

```{r rcell0}
library(dplyr)
library(glue)
```

## Real Case

It is the most common in the data manipulation that deriving a column based on a given formula, such as `Toxicity Grade variables`, and `PCSA variables`.

```{r rcell15}
#| echo: true
#| eval: true
df <- data.frame(SUBJID = paste0("S", 1:10),
                 GROUP = factor(sample(c("A", "B", "C"), 10, replace = T), levels = c("A", "B", "C")),
                 AVAL = sample.int(20, 10)
)
```

If we need to create new variables based on AVAL variable and condition or criteria, we have to do the following derivation.

```{r}
#| echo: true
#| eval: true
df <- df %>% mutate(level_1 = ifelse(AVAL>3 & GROUP == "A", "level 1", NA_character_),
                    level_2 = ifelse(AVAL<3, "level 2", NA_character_))
```
```{r rcell11612 }
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(df, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

we used the `ifelse` or `case_when` to implemente.In real case, the massive new variables and conditions lead to a not easy life!
here, we can develop functions for functional programmer purpose to handle this kind of task.

## Method 1

we store the column names, formula, and value in a list `derived_rule`.
```{r rcell116}
#| echo: true
#| eval: true

derived_rule <- list('level_1_new' = list('formula' = 'AVAL>3 & GROUP == "A"', value = "level 1"),
                'level_2_new' = list('formula' = 'AVAL<3', value = "level 2"))
```

we can combine the `glue` to build a function based on the given formula, and then make use of the `do.call` acheive our goal.

##### 1. convert formula to a function

```{r}
#| echo: true
#| eval: true
convert_to_function <- function(formula, value = NULL) {
  args <- paste(all.vars(rlang::parse_expr(formula)), collapse = ", ")
  args <- paste0(args, ", value = NULL")
  glue::glue(
    "function({args}) {{
        if(is.null(value)){{
            {formula}
        }}else{{
          ifelse({formula}, value, NA)
        }}
      }}"
  )
}
```
the involved columns in the formula will be used as the arguments in the new function.

##### 2.Build list with functions and arguments and label value

```{r}
#| echo: true
#| eval: true
derived_column_functions <- function(formula, value = NULL) {

  fun <- convert_to_function(formula)

  fun <- eval(rlang::parse_expr(fun))
  function_args <- rlang::fn_fmls_names(fun)
  function_args <- function_args[function_args!="value"]
  return(list(fun = fun,
              args = function_args,
              value = value))
}
```
For funtional programming, we use list to store the information of a new column, including the formula function, the needed arguments(columns), and the derived value.

##### 3.derive columns from formula

```{r}
#| echo: true
#| eval: true
derive_columns <- function(original_df, derive_infos) {
  df_names <- names(original_df)
  derived_names <- names(derive_infos)

  all_args <- unlist(purrr::map(derive_infos, ~.x$args))

  if (!all(all_args %in% c(df_names, derived_names))) {
    stop("derived columns require columns should exist in original data.")
  }

  columns <- recursive_column_calculation(original_df, derive_infos)

  return(columns)
}

# derive new columns
recursive_column_calculation <- function(original_df, derive_infos) {

    cols_name <- names(derive_infos)
    for (column_name in cols_name) {
        fun <- derive_infos[[column_name]]
        if (all(fun$args %in% names(original_df))) {
            original_df[[column_name]] <- do.call(fun$fun,
                                                    append(lapply(fun$args, \(x) original_df[[x]]), list(value = fun$value))) %>%
                as.vector()
            derive_infos[[column_name]] <- NULL
            if (length(derive_infos) == 0) {
                break()
            } else {
                original_df <- recursive_column_calculation(original_df, derive_infos)
            }
        }
    }
    return(original_df)
}

```
everthing is ready for deriving the new columns. we will find the argument `derive_infos` of `derive_columns` is a list that from the result of `derived_column_functions`.
that means however many columns there need to be derived, we can use the `purrr::map` to do a loop efficiently.

Based on the `deriven_rul` and original data, we can easily add new columns as below:

```{r}
#| echo: true
#| eval: true
derived_functions <- purrr::map(
  derived_rule,
  ~derived_column_functions(.x$formula, value = .x$value)
)
df1 <- derive_columns(df, derived_functions)
```
```{r}
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(df1, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

## Method 2

another way of handling the formula is evaluate the fomula string directly instead of converting it to function firstly.

```{r}
#| echo: true
#| eval: true
#| message: FALSE
cond <- c(
    quote(AVAL>3 & GROUP == "A")
)
df2 <- df %>% mutate(level_1 = ifelse(eval(do.call("substitute", cond)), 'level 1', NA_character_))
```
```{r}
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(df2, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

but I think it is neither robust enough on the perspective of data processing, nor the functional programming in R.
it is also easily lead to the risk of data leak in the process,  some records may can't be identified by this way.

### 2023-04-17 posts/manipulate/formula_derive.qmd
Update formula_derive.qmd

@@
 
-It is the most common in the data manipulation that deriving a column based on a given formula, such as `Toxicity Grade variables`, and `PCSA variables`.
+It is the most common case in the data manipulation that deriving a column based on a given formula, such as `Toxicity Grade variables`, and `PCSA variables`.
 
@@
 
-we used the `ifelse` or `case_when` to implemente.In real case, the massive new variables and conditions lead to a not easy life!
-here, we can develop functions for functional programmer purpose to handle this kind of task.
+we used the `ifelse` or `case_when` to implement the new variables.In real case, the massive new variables and conditions lead to a not easy life!
+here, we can develop functions for functional programming purpose to handle this kind of task.
 

### 2023-04-24 posts/manipulate/draft_tlf_generator.qmd
Create draft_tlf_generator.qmd

---
title: 'Draft TLF generator'
date: '2023-04-19'
author: Qing Zou
categories: ['R', 'R6class']
description: 'Draft'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

```{r rcell0}
library(checkmate)
library(R6)
library(assertive)
library(dplyr)
library(rlang)
library(Hmisc)
library(data.table)
box::use(R/utils)
```

## AdsModel

Create a `AdsModel` to manage the datasets level process

```{r rcell15}
#| echo: true
#| eval: true
#' @title AdsModel Class
#'
#'
#' @description
#' This is the  base class for ADaM dataset, to manage the CDISC attributes on dataset level
#'
#' AdsModel are build around the three following key parts:
#'
#' * Methods
#'
#'
#' For now, only implement that adding the attribute Model label by default. see `add_model_name()`
#'
#' test <- AdsModel$new()
#' test$initialize(adsl, adpsl)
#'
#' names(test$dataset)
#' a <- test$dataset
#'
#' tmp <- test$get_dataset('adsl')
#' attributes(tmp)
#'
#' adsl2 <- adsl
#' test$add_dataset(adsl2)
#' names(test$dataset)
#'
#' ```
#'

AdsModel <- R6::R6Class(
  "AdsModel",
  public = list(
    lib = NA,
    initialize = function(...){
      df_lst <-  tibble::lst(...)
      if(is.character(check_list(df_lst, min.len = 1))){
        # message("No datasets passed")
      }else{
        df_lst <- purrr::map2(df_lst, names(df_lst), ~utils$add_model_name(.x, .y))
        private$.dataset <- private$.dataset |> append(df_lst)
      }

    },
    get_dataset = function(domain){
      if(missing(domain)){
        return(private$.dataset)
      }else{
        checkmate::check_character(domain,
                                   min.chars = 1L,
                                   any.missing = FALSE)
        if(length(domain) == 1){
          return(purrr::keep_at(private$.dataset, domain)[[1]])
        }else{
          return(purrr::keep_at(private$.dataset, domain))
        }

      }
    },
    add_dataset = function(...){
        df_lst <-  tibble::lst(...)
        if(is.character(check_list(df_lst, min.len = 1))){
          message("No datasets passed")
        }else{
          df_lst <- purrr::map2(df_lst, names(df_lst), ~utils$add_model_name(.x, .y))
          private$.dataset <- private$.dataset |> append(df_lst)
        }

      },
    remove_dataset = function(...){
      # to be developed
    }
  ),
  active = list(
    dataset = function(){
        return(private$.dataset)
    }
  ),
  private = list(
    .dataset = list()
  )
)
```

```{r}
#| echo: true
#| eval: true
# test <- AdsModel$new(adsl)
adsl <- haven::read_xpt('../../data/adam/adsl.xpt')
test <- AdsModel$new()

test$initialize(adsl)

names(test$dataset)
```

```{r}
#| echo: true
#| eval: true

tmp <- test$get_dataset('adsl')
attributes(tmp)
```

```{r}
#| echo: true
#| eval: true

adtte <- haven::read_xpt('../../data/adam/adtte.xpt')
test$add_dataset(adtte)
names(test$dataset)
tmp2 <- test$get_dataset('adtte')
attributes(tmp2)
```

## VariableProc

Create a `VariableProc` to manage the variable level process

```{r rcell116}
#| echo: true
#| eval: true

VariableProc <- R6::R6Class(
  "VariableProc",
  inherit = AdsModel,
  public = list(
    # datasets level dictionary, can be retrieved by dataset
    # variables level dictionary, can be retrieved by dataset$variable
    dictionary = list(),
    initialize = function(dictionary_path = NA){

      #check dictionary_path is string
      if(!is.na(dictionary_path)){
        utils$check_dictionary_path(dictionary_path, "dictionary_path")
        dictionary <-  jsonlite::fromJSON(dictionary_path)
        self$dictionary <- dictionary
      }

    },
    get_decode = function(){
      if(is.character(check_list(self$dictionary, min.len = 1))){
        message("No dictionary avaiable, the variables will not be decoded")
      }else{
        private$.dataset <- purrr::map(private$.dataset, ~utils$set_attr(df = ., dictionary = self$dictionary))
      }

    },
    update_decode = function(dataset = NA, dictionary = ""){
      #check if the dataset exist first
      #check the format of dictionary

      # to be developed
    },
    update_detail = function(){
      # element track which dataset and variable be decoded, is it necessary integrated to S3 'summary'?

      # to be developed
    }
  ),

  #set dataset private, except variable attribute, user can't modify dataset in this module
  private = list(
    .dataset = list()
  )
)
```

```{r rcell11612 }
#| echo: true
#| eval: true
test2 <- VariableProc$new()

test2$initialize(dictionary_path = "dictionary.json")

test2$add_dataset(adsl)
adsl_tmp <- test2$get_dataset('adsl')

attributes(adsl_tmp)

attributes(adsl_tmp$SEX)$format

```

```{r}
test2$get_decode()

adsl_tmp2 <- test2$get_dataset('adsl')
attributes(adsl_tmp)
attributes(adsl_tmp2$SEX)$format

attributes(adsl_tmp2$AGEGR1)$format
```

### 2023-04-24 posts/manipulate/R/utils.R
Create utils.R

@@ -1,0 +1,59 @@
+#' Wrap the check for file exists
+#'
+#' @param x the object to be checked
+#' @param name the name of x
+#
+check_dictionary_path <- function(x, name) {
+    check_result_char <-
+        checkmate::check_character(x,
+                                   min.chars = 1L,
+                                   any.missing = FALSE,
+                                   len = 1L,
+                                   null.ok = FALSE
+        )
+
+    check_result_path <- checkmate::check_file_exists(x)
+
+    if (is.character(check_result_char)) {
+        stop(
+            stringr::str_glue("Argument '{name}' is invalid. {check_result_char}"),
+            call. = FALSE
+        )
+    } else if (is.character(check_result_path)) {
+        stop(
+            stringr::str_glue("Argument '{name}' is invalid. {check_result_path}"),
+            call. = FALSE
+        )
+    } else {
+        invisible(x)
+    }
+}
+
+
+#' add "format" attributes to the variables of passed datasets
+#'
+#' @param df
+#' @param dictionary
+#' @param .vars the variables used to be assisgned the attributes
+set_attr <- function(df, dictionary, .vars){
+
+    sub_dict <- purrr::keep_at(dictionary, colnames(df))
+
+    for (var in names(sub_dict)) {
+        var_N <- ifelse(paste0(var, 'N') %in% colnames(df), paste0(var, 'N'), 'NA')
+        attr(df[[var]], "format") <- list("Value" =  names(sub_dict[[var]]),
+                                          "Decode" = purrr::map_vec(sub_dict[[var]], 1),
+                                          "Numeric variable" = var_N,
+                                          "Ordinal" = names(sub_dict[[var_N]]))
+    }
+
+    return(df)
+}
+
+#' add a Model name attribute to data.frame
+add_model_name <- function(df, name){
+    attr(df, 'Model') <-toupper(name)
+    return(df)
+}
+
+

### 2023-04-26 posts/manipulate/draft_tlf_generator.qmd
Update draft_tlf_generator.qmd

@@
 title: 'Draft TLF generator'
-date: '2023-04-19'
+date: '2023-04-25'
 author: Qing Zou
 categories: ['R', 'R6class']
-description: 'Draft'
+description: 'Draft the TLF generator tool'
 execute:
@@
 library(data.table)
+library(rtables)
 box::use(R/utils)
+source("R/rtables_fct.R")
+source("R/tlf_pgm.R")
 ```
@@
 
+      },
+    load_dataset = function(path = NULL, domains){
+
+      df_lst <- lapply(domains, \(domain) haven::read_xpt(file.path(path, paste0(domain, ".xpt"))))
+      names(df_lst) <- domains
+      private$.dataset <- private$.dataset |> append(df_lst)
     },
@@
 ```
+
+### TLFProc
+
+Create a `TLFProc` to generate the outputs
+
+```{r}
+TLFProc <- R6::R6Class(
+  "TLFProc",
+  public = list(
+    tlfs = NULL,
+    tlfmetadata = NULL,
+    initialize = function(config = NULL) {
+      self$tlfs = new.env(parent = emptyenv())
+
+      if(!is.null(config)){
+        self$add_tlfmetadata(config)
+      }
+    },
+    has_outputs = function(pattern = NULL) {
+      outputs = ls(self$tlfs, all.names = TRUE)
+      if (!is.null(pattern)) {
+        assert_string(pattern)
+        outputs = outputs[grepl(pattern, outputs)]
+      }
+      outputs
+    },
+    get = function(output, ...) {
+      assert_string(output, min.chars = 1L)
+      # return output
+      obj = get0(output, envir = self$tlfs, inherits = FALSE, ifnotfound = NULL)
+      return(obj)
+    },
+    mget = function(outputs, ...) {
+      assert_character(outputs, min.chars = 1L, any.missing = FALSE)
+      # return
+      set_names(lapply(outputs, self$get, ...), outputs)
+    },
+    # read yaml configuration
+    add_tlfmetadata = function(config) {
+      assert_string(config, min.chars = 1L)
+      assert(check_file_exists(config))
+
+      metadata <- yaml::read_yaml(config)
+      self$tlfmetadata <- purrr::map(metadata, utils$parse_mtdata)
+      #
+    },
+
+    generator = function(VariableProc, tlfmetadata = self$tlfmetadata){
+
+      #check VariableProc
+      #check tlfmetadata
+
+      domains <- purrr::map(tlfmetadata, ~purrr::pluck(.x, "domains"))  |>
+        purrr::list_simplify()  |>
+        unique()
+
+      # dynamically construct a standard tlf template function with the metadata?
+      # args_list <- list()
+      purrr::walk2(tlfmetadata,
+                   names(tlfmetadata),
+                   ~assign(x = .y,
+                           value = do.call("demo_t_generator",
+                                           list(dataset = VariableProc$get_dataset(domains),
+                                                columns = .x$variables,
+                                                columns_label = .x$variables,
+                                                string_variable  = .x$string_variable,
+                                                header_variable = .x$header_variable,
+                                                header_order = .x$header_order,
+                                                population_variable = .x$population_variable)),
+                           envir = self$tlfs),
+                   .progress = TRUE)
+
+      invisible(self)
+    },
+    remove = function(keys) {
+      i = wf(!self$has(keys))
+      rm(list = keys, envir = self$items)
+      invisible(self)
+    }
+  )
+)
+```
+
+initialize a new instance of `TLFProc`, add tlfmetadata `mock.yaml`
+```{r}
+tlfp <- TLFProc$new()
+
+tlfp$add_tlfmetadata("mock.yaml")
+```
+
+<details><summary  class = "block2"> <font color="blue"> mock.yaml </font> __(click)__ </summary>
+
+```{r}
+#| echo: true
+#| eval: false
+
+demographic_example1:
+  columns:
+    AGEGR1:
+      domain: adsl
+      type: string
+      label: Age Goup (n[%])
+    AGE:
+      domain: adsl
+      type: numberic
+      label: Age
+      stats: N Mean Median Q1 Q3 Min Max
+    SEX:
+      domain: adsl
+      type: string
+      label: Sex (n[%])
+  header:
+    variable:
+      TRT01A
+    order:
+      TRT01AN
+  population:
+    variable:
+      SAFFL
+demographic_example2:
+  columns:
+    RACE:
+      domain: adsl
+      type: string
+      label: Race (n[%])
+    AGEGR1N:
+      domain: adsl
+      type: string
+      label: Age Goup (n[%])
+    BMIBL:
+      domain: adsl
+      type: numberic
+      label: Baseline BMI (kg/m^2)
+      stats: N Mean Median Q1 Q3 Min Max
+    BMIBLGR1:
+      domain: adsl
+      type: string
+      label: Baseline BMI Agroup
+  header:
+    variable:
+      SEX
+    order:
+      NULL
+  population:
+    variable:
+      DISCONFL
+```
+</details>
+
+<details><summary class = "block2">
+```{r}
+#| echo: true
+#| eval: false
+tlfp$tlfmetadata
+```
+</summary>
+```{r}
+#| echo: false
+#| eval: true
+tlfp$tlfmetadata
+```
+
+</details>
+
+Get all the needed datasets from metadata
+```{r}
+domains <- purrr::map(tlfp$tlfmetadata, ~purrr::pluck(.x, "domains"))  |>
+        purrr::list_simplify()  |>
+        unique()
+```
+
+- initialize a new instance of `VaribaleProc` with the dictionary of variables `dictionary.json`,
+- load needed  datasets
+- add attribute `format` to variables by `get_decode()`
+
+```{r}
+varpoc <- VariableProc$new()
+
+varpoc$initialize(dictionary_path = "dictionary.json")
+
+varpoc$load_dataset(path = "../../data/adam", domains = domains)
+varpoc$get_decode()
+```
+
+<details><summary class = "block2"> <font color="blue"> dictionary.json </font> __(click)__ </summary>
+
+```{r}
+#| echo: true
+#| eval: false
+{
+  "AGEGR1": {
+    "<65": ["<65"],
+    "65-80": ["65-80"],
+    ">80": [">80"]
+  },
+    "AGEGR1N": {
+    "1": ["<65"],
+    "2": ["65-80"],
+    "3": [">80"]
+  },
+  "SEX": {
+    "M": ["Male"],
+    "F": ["Female"]
+  },
+  "RACE": {
+    "WHITE": ["White"],
+    "BLACK OR AFRICAN AMERICAN": ["Black Or African American"],
+    "AMERICAN INDIAN OR ALASKA NATIVE": ["American Indian Or Alaska Native"]
+  },
+  "BMIBLGR1": {
+    "<25": ["<25"],
+    "25-<30": ["25-<30"],
+    ">=30": [">=30"]
+  }
+}
+```
+</details>
+
+- Generate the outputs
+```{r}
+tlfp$generator(VariableProc = varpoc)
+```
+
+- Check what the outputs we have in `tlfp`
+```{r}
+tlfp$has_outputs()
+
+tlfp$has_outputs("1")
+```
+
+- the tlf metadata of the table `demographic_example1`
+```{r}
+tlfp$tlfmetadata$demographic_example1
+```
+
+```{r}
+#| echo: false
+#| eval: true
+
+#' Display rtables in html
+#'
+#' @param table a rtables object
+#' @param to_html export to HTML. Default TRUE
+#'
+#' @return a html table
+#' @export
+table_to_html <- function(table, to_html = TRUE) {
+  stopifnot(!is.null(table), class(table) %in% c("TableTree", "ElementaryTable"))
+  extra_css <- "
+  .tabwid {overflow-x: scroll; display: block;}
+  .tabwid table td p {min-width: 170px !important;}
+  .tabwid table td:first-child > p {min-width: 200px !important;}
+  .tabwid table tr:hover {background-color: #ffff99 !important;}
+  "
+  flextable::set_flextable_defaults(
+    font.family = ifelse(to_html, "Source Sans Pro", "Verdana"),
+    extra_css = extra_css
+  )
+  ft <- rtables::tt_to_flextable(table) %>%
+    flextable::border_remove() %>%
+    flextable::bold(bold = TRUE, part = "header") %>%
+    flextable::hline_top(part = "body") %>%
+    flextable::hline_bottom(part = "body") %>%
+    flextable::autofit()
+  if (to_html) {
+    flextable::htmltools_value(ft)
+  } else {
+    ft
+  }
+}
+```
+
+- Retrieve the outputs
+```{r}
+#| echo: true
+#| eval: false
+tlfp$get("demographic_example1")
+```
+```{r}
+#| echo: false
+#| eval: true
+#| message: FALSE
+table_to_html(tlfp$get("demographic_example1"), to_html = FALSE)
+```
+
+```{r}
+
+tlfp$tlfmetadata$demographic_example1
+
+```
+
+```{r}
+#| echo: true
+#| eval: false
+tlfp$get("demographic_example2")
+```
+```{r}
+#| echo: false
+#| eval: true
+#| message: FALSE
+table_to_html(tlfp$get("demographic_example2"), to_html = FALSE)
+```
+
+the size of tlfp
+```{r}
+print(object.size(tlfp))
+```
+
+Transform the class of `rtables`
+```{r}
+table_to_dataframe <- function(tbl) {
+
+  # check the tabl classs
+
+  matrix_tbl <- tbl %>% rtables::matrix_form()
+
+  tbl_strings <- as.data.frame(matrix_tbl$strings)
+
+  ## get the hader index from tbl attributes
+  nrow_header <- attributes(matrix_tbl)$nrow_header
+
+  ## get the body of tbl
+  tbl_body <- tbl_strings[-c(1:nrow_header), ]
+
+  ## get the header of tbl
+  tbl_header <- tbl_strings[c(1:nrow_header), ]
+
+  attr(tbl_body, "tbl_header") <- tbl_header
+
+  return(tbl_body)
+}
+```
+
+```{r}
+tbl <-  tlfp$get("demographic_example2") |> table_to_dataframe()
+```
+```{r du2 }
+#| echo: false
+#| eval: true
+#| message: FALSE
+knitr::kable(tbl, "html")%>%
+  kableExtra::kable_styling(font_size = 16)
+```
+```{r}
+attributes(tbl)
+```

### 2023-04-26 posts/manipulate/R/utils.R
Update utils.R

@@ -57,3 +57,23 @@ add_model_name <- function(df, name){
 }
 
 
+#######TLF----------------
+
+parse_mtdata <- function(metadata){
+
+    # check the structure of metadata
+
+    #------------------------------#
+    #------------------------------#
+
+
+    #check the results
+
+    return(list(variables = names(metadata$columns),
+                string_variable = names(purrr::keep(metadata$columns, ~.x$type == "string")),
+                domains = purrr::map_vec(metadata$columns, ~purrr::pluck(.x, 'domain'))  |>  unique(),
+                header_variable = purrr::pluck(metadata, "header")$variable,
+                header_order = purrr::pluck(metadata, "header")$order,
+                population_variable = purrr::pluck(metadata, "population")$variable)
+    )
+}

### 2023-04-26 posts/manipulate/R/tlf_pgm.R
Create tlf_pgm.R

@@ -1,0 +1,48 @@
+
+demo_t_generator <- function(dataset, columns, columns_label, string_variable, header_variable, header_order, population_variable){
+
+
+    names(header_variable) <- NULL
+
+    adsl <- dataset %>%
+        select(USUBJID, all_of(population_variable), all_of(header_variable), all_of(header_order), all_of(columns)) %>%
+        filter(!!as.name(population_variable) == "Y")
+    # %>%
+    #   mutate(across(string_variable, ~ifelse(.=="", "Missing_value_impute", as.character(.))))
+
+    adsl_fact <- adsl %>%
+        mutate(across(all_of(string_variable),
+                      ~factor(attributes(cur_data()[[cur_column()]])$format$Decode[.x],
+                              levels = attributes(cur_data()[[cur_column()]])$format$Decode)))
+
+    if(isTRUE(is.na(header_order))){
+        adsl_fact <- adsl_fact %>%
+            mutate(across(all_of(header_variable),
+                          ~factor(attributes(cur_data()[[cur_column()]])$format$Decode[.x],
+                                  levels = attributes(cur_data()[[cur_column()]])$format$Decode)))
+    }else{
+        adsl_fact <- df_order_header(adsl_fact, c(header_variable, header_order))
+    }
+
+
+
+    col_counts <- table(adsl[[header_variable]]) %>% as.vector()
+    col_counts <- c(col_counts, sum(col_counts))
+
+    calculated_t <- basic_table() %>%
+        split_cols_by(header_variable) %>%
+        add_overall_col("All") %>%
+        summarize_vars(
+            vars = columns,
+            var_labels = columns_label,
+            dispfmt = c("npct"),
+            decimal = 1,
+            denom = c("col"),
+            unique_row = TRUE
+        ) %>%
+        rtables::build_table(adsl_fact, col_counts = col_counts)
+
+    return(calculated_t)
+}
+
+

### 2023-04-26 posts/manipulate/R/rtables_fct.R
Create rtables_fct.R

@@ -1,0 +1,295 @@
+
+#' a pair of functions sprintf_format and format_values used for rcell
+#'
+#' @param n Unique number of events for the current category
+#' @param N Total number of events under the split column
+#' @param decimal Number of decimal to display
+#' @param type to decide if the denominator n/N(%) in the result
+#' @keywords sprintf_format
+#'
+#' @return
+sprintf_format_helper <- function(n, N, decimal = 1, type = "n(%)") {
+    showN <- NULL
+    if (type == "n/N(%)"){
+        showN <- "/%.0f"
+    }
+
+    if (n == 0) {
+        fmt <- sprintf_format(paste0("%.0f", showN))
+    } else if (n == N) {
+        # if pct == 100%, then show (100) instead of (100.0)
+        fmt <- sprintf_format(paste0("%.0f", showN, " ", "(%.0f)"))
+    } else {
+        fmt <- sprintf_format(paste0("%.0f", showN, " (%.", decimal, "f)"))
+    }
+
+    return(fmt)
+}
+
+
+#' a pair of functions sprintf_format and format_values used for rcell
+#'
+#' @param n Unique number of events for the current category
+#' @param N Total number of events under the split column
+#' @param type to decide if the denominator n/N(%) in the result
+#'
+#' @keywords format_values
+#' @example
+#' rcell(x = format_values_helper(1, 10),
+#'       format = sprintf_format_helper(1, 10))
+#'
+#' rcell(x = format_values_helper(1, 10, type = "n/N(%)"),
+#'       format = sprintf_format_helper(1, 10, type = "n/N(%)"))
+#'
+#' rcell(x = format_values_helper(10, 10),
+#'       format = sprintf_format_helper(10, 10))
+#'
+#' @return
+format_values_helper <- function(n, N, type = "n(%)") {
+    showN <- NULL
+    if (type == "n/N(%)"){
+        showN <- N
+    }
+    if (n == 0) {
+        value <- c(0, showN)
+    } else {
+        value <- c(n, showN, n / N * 100)
+    }
+
+    return(value)
+}
+
+
+df_order_header <- function(df, header_var, sort_df = FALSE, add_all = FALSE) {
+    if (is.null(df)) {
+        stop("missing df")
+    }
+
+    if (is.null(header_var)) {
+        stop("missing header_var")
+    }
+
+    # settings
+    var_name <- header_var[1]
[…]

### 2023-04-26 posts/manipulate/draft_tlf_generator.qmd
Update draft_tlf_generator.qmd

@@
 
-Get all the needed datasets from metadata
+- Get all the needed datasets from metadata
 ```{r}
@@
 ```
+&nbsp;
 
-the size of tlfp
+- the size of tlfp
 ```{r}
@@
 
-Transform the class of `rtables`
+- Transform the class of `rtables`
 ```{r}

### 2023-04-26 posts/manipulate/draft_tlf_generator.qmd
Update draft_tlf_generator.qmd

@@
 author: Qing Zou
-categories: ['R', 'R6class']
+categories: ['R', 'R6class', 'TLF']
 description: 'Draft the TLF generator tool'
@@
 #'
-#' For now, only implement that adding the attribute Model label by default. see `add_model_name()`
-#'
-#' test <- AdsModel$new()
-#' test$initialize(adsl, adpsl)
-#'
-#' names(test$dataset)
-#' a <- test$dataset
-#'
-#' tmp <- test$get_dataset('adsl')
-#' attributes(tmp)
-#'
-#' adsl2 <- adsl
-#' test$add_dataset(adsl2)
-#' names(test$dataset)
-#'
-#' ```
-#'
 
@@
     public = list(
-    lib = NA,
         initialize = function(...){
@@
         },
+        #' retrieve datasets from AdsModel by names
         get_dataset = function(domain){
@@
         },
+        #' load dataset from disk
+        #'
+        #' @param path the database directory
+        #' @param domain the dataset names
         load_dataset = function(path = NULL, domains){
@@
 
+#### Used functions
+
+<details><summary  class = "block2"> <font color="blue"> add_model_name </font> __(click)__ </summary>
+```{r}
+#| echo: true
+#| eval: false
+#' add a Model name attribute to data.frame
+add_model_name <- function(df, name){
+    attr(df, 'Model') <-toupper(name)
+    return(df)
+}
+```
+</details>
+
+#### Steps
 ```{r}
@@
 
+# retrieve data by 'dataset'
 names(test$dataset)
@@
 tmp <- test$get_dataset('adsl')
-attributes(tmp)
+#the updated attribute 'Model'
+attributes(tmp)$Model
 ```
@@
 adtte <- haven::read_xpt('../../data/adam/adtte.xpt')
+
+# add dataset to instance
 test$add_dataset(adtte)
@@
 tmp2 <- test$get_dataset('adtte')
-attributes(tmp2)
+attributes(tmp2)$Model
 ```
@@
   public = list(
-    # datasets level dictionary, can be retrieved by dataset
-    # variables level dictionary, can be retrieved by dataset$variable
+    # save the dictionary
     dictionary = list(),
@@
 
+- the `VariableProc` inherit from `AdsModel`
+- method `get_decode` set attributes format to variables
+
+#### Used functions
+
+<details><summary  class = "block2"> <font color="blue"> set_attr </font> __(click)__ </summary>
+```{r}
+#| echo: true
+#| eval: false
+#' add "format" attributes to the variables of passed datasets
+#'
+#' @param df
+#' @param dictionary
+#' @param .vars the variables used to be assisgned the attributes
+set_attr <- function(df, dictionary, .vars){
+
+    sub_dict <- purrr::keep_at(dictionary, colnames(df))
+
+    for (var in names(sub_dict)) {
+        var_N <- ifelse(paste0(var, 'N') %in% colnames(df), paste0(var, 'N'), 'NA')
+        attr(df[[var]], "format") <- list("Value" =  names(sub_dict[[var]]),
+                                          "Decode" = purrr::map_vec(sub_dict[[var]], 1),
+                                          "Numeric variable" = var_N,
+                                          "Ordinal" = names(sub_dict[[var_N]]))
+    }
+
+    return(df)
+}
+
+```
+</details>
+
+#### Steps
+
 ```{r rcell11612 }
@@
 
-attributes(adsl_tmp)
+# inherit the AdsModel method
+attributes(adsl_tmp)$Model
 
+# before setting the format
 attributes(adsl_tmp$SEX)$format
@@
 ```{r}
+
+# set the format by get_decode()
 test2$get_decode()
@@
 adsl_tmp2 <- test2$get_dataset('adsl')
-attributes(adsl_tmp)
+
+# the format is set from dictionary.json
 attributes(adsl_tmp2$SEX)$format
@@
   public = list(
+
+    #' Stores the object of the outputs
     tlfs = NULL,
+
+    #' Store the metadata
     tlfmetadata = NULL,
+
+    #' create a new environment for TLFProc
     initialize = function(config = NULL) {
       self$tlfs = new.env(parent = emptyenv())
-
       if(!is.null(config)){
@@
     },
+    #' return avaialable outputs name
     has_outputs = function(pattern = NULL) {
@@
     },
+
+    #' get output by name
     get = function(output, ...) {
@@
     },
+
+    #' get a list of outputs
     mget = function(outputs, ...) {
@@
 
+    #' Generate the TLFs
+    #' @param VariableProc the instance VariableProc with decoded dataset
+    #' @param tlmetadata  the passed metadata
     generator = function(VariableProc, tlfmetadata = self$tlfmetadata){
@@
 ```
+#### Used functions
 
-initialize a new instance of `TLFProc`, add tlfmetadata `mock.yaml`
+<details><summary  class = "block2"> <font color="blue"> parse_mtdata </font> __(click)__ </summary>
+```{r}
+#| echo: true
+#| eval: false
+parse_mtdata <- function(metadata){
+
+    # check the structure of metadata
+
+    #------------------------------#
+    #------------------------------#
+
+    #check the results
+
+    return(list(variables = names(metadata$columns),
+                string_variable = names(purrr::keep(metadata$columns, ~.x$type == "string")),
+                domains = purrr::map_vec(metadata$columns, ~purrr::pluck(.x, 'domain'))  |>  unique(),
+                header_variable = purrr::pluck(metadata, "header")$variable,
+                header_order = purrr::pluck(metadata, "header")$order,
+                population_variable = purrr::pluck(metadata, "population")$variable)
+    )
+}
+```
+</details>
+
+<details><summary  class = "block2"> <font color="blue"> demo_t_generator </font> __(click)__ </summary>
+```{r}
+#| echo: true
+#| eval: false
+
+demo_t_generator <- function(dataset, columns, columns_label, string_variable, header_variable, header_order, population_variable){
+
+    names(header_variable) <- NULL
+
+    adsl <- dataset %>%
+        select(USUBJID, all_of(population_variable), all_of(header_variable), all_of(header_order), all_of(columns)) %>%
+        filter(!!as.name(population_variable) == "Y")
+    # %>%
+    #   mutate(across(string_variable, ~ifelse(.=="", "Missing_value_impute", as.character(.))))
+
+    adsl_fact <- adsl %>%
+        mutate(across(all_of(string_variable),
+                      ~factor(attributes(cur_data()[[cur_column()]])$format$Decode[.x],
+                              levels = attributes(cur_data()[[cur_column()]])$format$Decode)))
+
+    if(isTRUE(is.na(header_order))){
+        adsl_fact <- adsl_fact %>%
+            mutate(across(all_of(header_variable),
+                          ~factor(attributes(cur_data()[[cur_column()]])$format$Decode[.x],
+                                  levels = attributes(cur_data()[[cur_column()]])$format$Decode)))
+    }else{
+        adsl_fact <- df_order_header(adsl_fact, c(header_variable, header_order))
+    }
+
+    col_counts <- table(adsl[[header_variable]]) %>% as.vector()
+    col_counts <- c(col_counts, sum(col_counts))
+
+    calculated_t <- basic_table() %>%
+        split_cols_by(header_variable) %>%
+        add_overall_col("All") %>%
+        summarize_vars(
+            vars = columns,
+            var_labels = columns_label,
+            dispfmt = c("npct"),
+            decimal = 1,
+            denom = c("col"),
+            unique_row = TRUE
+        ) %>%
+        rtables::build_table(adsl_fact, col_counts = col_counts)
+
+    return(calculated_t)
+}
+
+```
+</details>
+
+#### Steps
+- initialize a new instance of `TLFProc`, add tlfmetadata `mock.yaml`
 ```{r}
@@
 
-<details><summary  class = "block2"> <font color="blue"> mock.yaml </font> __(click)__ </summary>
+<details><summary  class = "block2"> <font color="blue"> demo_t_generator </font> __(click)__ </summary>
 

### 2023-04-27 posts/manipulate/draft_tlf_generator.qmd
Update draft_tlf_generator.qmd

@@
 ---
+
+Please find the code of this section in <font color="blue"> [add_model_name](https://posit.cloud/content/5870060) </font>
 
@@
 
-<details><summary  class = "block2"> <font color="blue"> demo_t_generator </font> __(click)__ </summary>
+<details><summary  class = "block2"> <font color="blue"> mock.yaml </font> __(click)__ </summary>
 

### 2023-04-28 posts/manipulate/draft_tlf_generator.qmd
Update draft_tlf_generator.qmd

@@
 
-Please find the code of this section in <font color="blue"> [add_model_name](https://posit.cloud/content/5870060) </font>
+Please find the code of this section in <font color="blue"> https://posit.cloud/content/5870060 </font>
 
@@
     },
-    remove = function(keys) {
-      i = wf(!self$has(keys))
-      rm(list = keys, envir = self$items)
+    remove = function(outputs) {
+      #check the outputs if exsit firstly
+      rm(list = outputs, envir = self$tlfs)
       invisible(self)

### 2023-04-28 posts/manipulate/draft_tlf_generator.qmd
Update draft_tlf_generator.qmd

@@
       #check the outputs if exsit firstly
+      #checkmate::wf(exsit(outputs, envir = self$tlfs))?
       rm(list = outputs, envir = self$tlfs)

### 2023-05-12 posts/report/data.validator.qmd
Add files via upload

---
title: 'Create a QC validation report by package data.validator'
date: '2023-05-12'
author: Qing Zou
categories: ['report', 'data.validator']
description: 'way to automated validation'
execute:
  message: false
  warning: false
editor_options:
  chunk_output_type: console
---

In the `TLFgenerator`, we get two demographic example 1 and 2.

```{r lib}
#| echo: false
#| eval: true
library(dplyr)
```

```{r read1}
#| echo: true
#| eval: true
example1 <- readRDS('data/example1.rds')
```

```{r ex1}
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(example1, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

```{r read2}
#| echo: true
#| eval: true
example2 <- readRDS('data/example2.rds')
```

```{r}
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(example2, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

to present the feature of data.validator, we change some value of the example output:

```{r}
#| echo: true
#| eval: true
example1_qc <- example1
example1_qc[3,1] <- "Age Group [n %]"
```

```{r}
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(example1_qc, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

```{r}
#| echo: true
#| eval: true
example2_qc <- example2
example2_qc[23, 1] <- "^{unicode} 30"
example2_qc[18, 2] <- "15 ; 38"
```

```{r}
#| echo: false
#| eval: true
#| message: FALSE
knitr::kable(example2_qc, "html")%>%
  kableExtra::kable_styling(font_size = 16)
```

```{r}
#| echo: true
#| eval: true
#| message: FALSE

library(assertr)
library(data.validator)

check_cols_value <- function(expected_value){
  function(x) { x %in% expected_value}
}

report <- data_validation_report()

validate(example1, name = "Verifying demographic_example1") %>%
  validate_cols(check_cols_value(example1_qc$V1), V1, description = "Check the first column V1") %>% 
  validate_cols(check_cols_value(example1_qc$V2), V2, description = "Check the first column V2") %>% 
  validate_cols(check_cols_value(example1_qc$V3), V3, description = "Check the first column V3") %>% 
  validate_cols(check_cols_value(example1_qc$V4), V4, description = "Check the first column V4") %>% 
  add_results(report)

validate(example2, name = "Verifying demographic_example2") %>%
  validate_cols(check_cols_value(example2_qc$V1), V1, description = "Check the first column V1") %>% 
  validate_cols(check_cols_value(example2_qc$V2), V2, description = "Check the first column V2") %>% 
  validate_cols(check_cols_value(example2_qc$V3), V3, description = "Check the first column V3") %>% 
  validate_cols(check_cols_value(example2_qc$V4), V4, description = "Check the first column V4") %>% 
  add_results(report)

```

```{r}
#| echo: true
#| eval: false
save_report(report, output_file = "Check_TLFs.html", output_dir = getwd())
```
<iframe src="Check_TLFs.html" height="100%" width="100%"></iframe>
